### 2014-03-28 tests/tests.Rmd

##Some other attempts

We try dividing the data by the standard deviation so that the new data will be iid. Here we work with the true standard deviation.

```{r}
source(file.path("../Rcode/bayesmooth.R"))
spike.f=function(x) (0.75*exp(-500*(x-0.23)^2)+1.5*exp(-2000*(x-0.33)^2)+3*exp(-8000*(x-0.47)^2)+2.25*exp(-16000*(x-0.69)^2)+0.5*exp(-32000*(x-0.83)^2))

pos = c(.1, .13, .15, .23, .25, .40, .44, .65, .76, .78, .81)
hgt = 2.97/5*c(4, 5, 3, 4, 5, 4.2, 2.1, 4.3, 3.1, 5.1, 4.2)
wth = c(.005, .005, .006, .01, .01, .03, .01, .01, .005, .008, .005)
mu.b = rep(0,n)
for(j in 1:length(pos)){
  mu.b = mu.b + hgt[j]/(( 1 + (abs(t - pos[j])/wth[j]))^4)
}

n=1024
t=1:n/n

mu.s=spike.f(t)

mu.t=(1+mu.s)/5

rsnr=sqrt(3)

var1=rep(1,n)
var2=(0.0001+4*(exp(-550*(t-0.2)^2)+exp(-200*(t-0.5)^2)+exp(-950*(t-0.8)^2)))/1.35
var3=0.00001+mu.b
sigma.ini=sqrt(var3)
sigma.t=sigma.ini/mean(sigma.ini)*sd(mu.t)/rsnr^2

set.seed(1025)
X.s=rnorm(n,mu.t,sigma.t)

mu.est<-bayesmooth(X.s,sigma=sigma.t)
mu.est.sig=sigma.t*bayesmooth(X.s/sigma.t,sigma=rep(1,n))

mse(mu.est,mu.t)
mse(mu.est.sig,mu.t)
```

```{r fig.width=12, fig.height=12}
plot(mu.t,type='l')
lines(mu.est,col=2)
lines(mu.est.sig,col=4)

```

So dividing the data by its SD does change some of the underlying properties of the original intensity, and does not perform as well as one hopes.

Next we look at better ways to estimate the variance. One way would be to log transform the squared residuals so that they have constant variance. We also try using jash and vash instead of ash. For simplicity, we assume that the true mean is known.

```{r}

spike.f=function(x) (0.75*exp(-500*(x-0.23)^2)+1.5*exp(-2000*(x-0.33)^2)+3*exp(-8000*(x-0.47)^2)+2.25*exp(-16000*(x-0.69)^2)+0.5*exp(-32000*(x-0.83)^2))
n=2^12
t=1:n/n
mu.s=spike.f(t)
pos = c(.1, .13, .15, .23, .25, .40, .44, .65, .76, .78, .81)
hgt = c(2, 3.6, 2, 5, 7.5, 6.9, 2, 4.8, 2, 4.1, 2.3)
wth = c(.005, .005, .006, .01, .01, .03, .01, .01, .005, .008, .005)
mu.b = rep(0,n)
for(j in 1:length(pos)){
  mu.b = mu.b + hgt[j]/(( 1 + (abs(t - pos[j])/wth[j]))^4)
}
dop.f=function(x) sqrt(x*(1-x))*sin((2*pi*1.05)/(x+0.05))
mu.dop=dop.f(t)
mu.sine=sin(20*t)
var1=1.42*((3-20*t)*(t>=0&t<0.1)+(20*t-1)*(t>=0.1&t<0.25)+(4+(1-4*t)*18/19)*(t>=0.25&t<0.725)+(2.2+10*(t-0.725))*(t>=0.725&t<0.89)+(3.85-85*(t-0.89)/11)*(t>=0.89&t<=1))
var2=(1+4*(exp(-550*(t-0.2)^2)+exp(-200*(t-0.5)^2)+exp(-950*(t-0.8)^2)))/1.35
var3=mu.b
var4=3.4*(2+mu.dop)
sigma.t.1=sqrt(var1)
sigma.t.2=sqrt(var2)
sigma.t.3=sqrt(var3)
sigma.t.4=sqrt(var4)
mu.t.1=rep(0,n)
mu.t.2=mu.sine

l2norm=function(x) sum(x^2)
mise=function(x,y) l2norm(x-y)/l2norm(y)
mse=function(x,y) mean((x-y)^2)

set.seed(327)
X.11=rnorm(n,mu.t.1,sigma.t.1)
set.seed(327)
X.12=rnorm(n,mu.t.1,sigma.t.2)
set.seed(327)
X.13=rnorm(n,mu.t.1,sigma.t.3)
set.seed(327)
X.14=rnorm(n,mu.t.1,sigma.t.4)
set.seed(327)
X.21=rnorm(n,mu.t.2,sigma.t.1)
set.seed(327)
X.22=rnorm(n,mu.t.2,sigma.t.2)
set.seed(327)
X.23=rnorm(n,mu.t.2,sigma.t.3)
set.seed(327)
X.24=rnorm(n,mu.t.2,sigma.t.4)

source("../simulation_1d_g/wip/bayesmooth_cur.R")
system.time(var.est.cur.11<-bayesmooth.cur(X.11,mu.t.1))
system.time(var.est.cur.12<-bayesmooth.cur(X.12,mu.t.1))
system.time(var.est.cur.13<-bayesmooth.cur(X.13,mu.t.1))
system.time(var.est.cur.14<-bayesmooth.cur(X.14,mu.t.1))
system.time(var.est.cur.21<-bayesmooth.cur(X.21,mu.t.2))
system.time(var.est.cur.22<-bayesmooth.cur(X.22,mu.t.2))
system.time(var.est.cur.23<-bayesmooth.cur(X.23,mu.t.2))
system.time(var.est.cur.24<-bayesmooth.cur(X.24,mu.t.2))

source("../simulation_1d_g/wip/bayesmooth_log.R")
system.time(var.est.log.11<-bayesmooth.log(X.11,mu.t.1,sigma.t.1))
system.time(var.est.log.12<-bayesmooth.log(X.12,mu.t.1,sigma.t.2))
system.time(var.est.log.13<-bayesmooth.log(X.13,mu.t.1,sigma.t.3))
system.time(var.est.log.14<-bayesmooth.log(X.14,mu.t.1,sigma.t.4))
system.time(var.est.log.21<-bayesmooth.log(X.21,mu.t.2,sigma.t.1))
system.time(var.est.log.22<-bayesmooth.log(X.22,mu.t.2,sigma.t.2))
system.time(var.est.log.23<-bayesmooth.log(X.23,mu.t.2,sigma.t.3))
system.time(var.est.log.24<-bayesmooth.log(X.24,mu.t.2,sigma.t.4))

source("../simulation_1d_g/wip/bayesmooth_vash.R")
system.time(var.est.vash.11<-bayesmooth.vash(X.11,mu.t.1))
system.time(var.est.vash.12<-bayesmooth.vash(X.12,mu.t.1))
system.time(var.est.vash.13<-bayesmooth.vash(X.13,mu.t.1))
system.time(var.est.vash.14<-bayesmooth.vash(X.14,mu.t.1))
system.time(var.est.vash.21<-bayesmooth.vash(X.21,mu.t.2))
system.time(var.est.vash.22<-bayesmooth.vash(X.22,mu.t.2))
system.time(var.est.vash.23<-bayesmooth.vash(X.23,mu.t.2))
system.time(var.est.vash.24<-bayesmooth.vash(X.24,mu.t.2))

source("../simulation_1d_g/wip/bayesmooth_jash.R")
system.time(var.est.jash.11<-bayesmooth.jash(X.11,mu.t.1))
system.time(var.est.jash.12<-bayesmooth.jash(X.12,mu.t.1))
system.time(var.est.jash.13<-bayesmooth.jash(X.13,mu.t.1))
system.time(var.est.jash.14<-bayesmooth.jash(X.14,mu.t.1))
system.time(var.est.jash.21<-bayesmooth.jash(X.21,mu.t.2))
system.time(var.est.jash.22<-bayesmooth.jash(X.22,mu.t.2))
system.time(var.est.jash.23<-bayesmooth.jash(X.23,mu.t.2))
system.time(var.est.jash.24<-bayesmooth.jash(X.24,mu.t.2))

matrix(c(mse(var.est.cur.11,sigma.t.1^2),
mse(var.est.cur.12,sigma.t.2^2),
mse(var.est.cur.13,sigma.t.3^2),
mse(var.est.cur.14,sigma.t.4^2),
mse(var.est.log.11,sigma.t.1^2),
mse(var.est.log.12,sigma.t.2^2),
mse(var.est.log.13,sigma.t.3^2),
mse(var.est.log.14,sigma.t.4^2),
mse(var.est.vash.11,sigma.t.1^2),
mse(var.est.vash.12,sigma.t.2^2),
mse(var.est.vash.13,sigma.t.3^2),
mse(var.est.vash.14,sigma.t.4^2),
mse(var.est.jash.11,sigma.t.1^2),
mse(var.est.jash.12,sigma.t.2^2),
mse(var.est.jash.13,sigma.t.3^2),
mse(var.est.jash.14,sigma.t.4^2)),
4,4)

matrix(c(mse(var.est.cur.21,sigma.t.1^2),
mse(var.est.cur.22,sigma.t.2^2),
mse(var.est.cur.23,sigma.t.3^2),
mse(var.est.cur.24,sigma.t.4^2),
mse(var.est.log.21,sigma.t.1^2),
mse(var.est.log.22,sigma.t.2^2),
mse(var.est.log.23,sigma.t.3^2),
mse(var.est.log.24,sigma.t.4^2),
mse(var.est.vash.21,sigma.t.1^2),
mse(var.est.vash.22,sigma.t.2^2),
mse(var.est.vash.23,sigma.t.3^2),
mse(var.est.vash.24,sigma.t.4^2),
mse(var.est.jash.21,sigma.t.1^2),
mse(var.est.jash.22,sigma.t.2^2),
mse(var.est.jash.23,sigma.t.3^2),
mse(var.est.jash.24,sigma.t.4^2)),
4,4)

```