### 2018-08-23 analysis/spikesdemo.Rmd
wflow_publish("spikesdemo.Rmd")

---
title: "Spikes demo"
author: "Zhengrong Xing, Peter Carbonetto and Matthew Stephens"
output: workflowr::wflow_html
---

This script illustrates key features of the SMASH method on a small,
simulated data set with heteroskedastic variances.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,fig.align = "center")
```

Set up environment
------------------

We begin by loading the ashr, smashr, scales, ggplot2 and cowplot
packages, as well as some additional functions used in the code below.

```{r load-pkgs, warning=FALSE}
library(ashr)
library(smashr)
library(scales)
library(ggplot2)
suppressMessages(library(cowplot))
source("../code/misc.R")
```

Define the "spikes" mean function
---------------------------------

Variable `mu.sp` defines the mean signal.

```{r define-spikes}
n <- 1024
t <- 1:n/n
spike.f <-
  function (x) (0.75 * exp(-500   * (x - 0.23)^2) +
                1.5  * exp(-2000  * (x - 0.33)^2) +
                3    * exp(-8000  * (x - 0.47)^2) +
                2.25 * exp(-16000 * (x - 0.69)^2) +
                0.5  * exp(-32000 * (x - 0.83)^2))
mu.sp <- spike.f(t)
mu.sp <- (1 + mu.sp)/5
```

Simulate data
-------------

Simulate the noisy observations from the normal distribution with
means that are given by the "spikes" function, and with
spatially-structured variances.

```{r simulate-data}
pos <- c(0.1,0.13,0.15,0.23,0.25,0.4,0.44,0.65,0.76,0.78,0.81)
hgt <- 2.88/5 * c(4,-5,3,-4,5,-4.2,2.1,4.3,-3.1,2.1,-4.2)
sig.cb <- rep(0,length(t))
for (j in 1:length(pos)) 
  sig.cb <- sig.cb + (1 + sign(t - pos[j])) * (hgt[j]/2)
sig.cb[sig.cb < 0] <- 0
sig.cb <- 0.1 + (sig.cb - min(sig.cb))/max(sig.cb)
rsnr   <- sqrt(3)
sig.cb <- sig.cb/mean(sig.cb) * sd(mu.sp)/rsnr^2
x.sim  <- rnorm(n,mu.sp,sig.cb)
```

Plot the simulated data set.

```{r plot-simulated-data, fig.height=5, fig.width=10}
par(cex.axis = 1,cex.lab = 1.5)
plot(mu.sp,type = 'l',ylim = c(-0.05,1),xlab = "position",
     ylab = "",lwd = 1.7,xaxp = c(0,1024,4),yaxp = c(0,1,4))
lines(mu.sp + 2*sig.cb,col = "darkorange",lty = 5,lwd = 1.8)
lines(mu.sp - 2*sig.cb,col = "darkorange",lty = 5,lwd = 1.8)
points(x.sim,cex = 0.7,pch = 16,col = "darkblue")
```

This plot shows the "spikes" mean function (the black line),
+/- 2 standard deviations (the yellow lines), and
the simulated data (the blue points).

Run SMASH and TI thresholding
-----------------------------

Now we apply SMASH, as well as the translation invariant (TI)
thresholding method, to the "spikes" data. Here we run the TI
thresholding method twice---once when the variance is assumes to be
constant ("homoskedastic"), and once when it is estimated using the
RMAD algorithm. The first method yields very poor estimates, so they
are not shown in the plots below.

The SMASH method is able to estimate the variance.

```{r denoise-signals}
sig.est  <- sqrt(2/(3 * (n - 2)) *
              sum((1/2 * x.sim[1:(n-2)] - x.sim[2:(n-1)] + x.sim[3:n])^2/2))
mu.smash <- smash(x.sim,family = "DaubLeAsymm",filter.number = 8)
mu.ti.homo <- ti.thresh(x.sim,sigma = sig.est,family = "DaubLeAsymm",
                        filter.number = 8)
mu.ti    <- ti.thresh(x.sim,method = "rmad",family = "DaubLeAsymm",
                      filter.number = 8)
```

It should take at the very most a few minutes to run SMASH and TI
thresholding.

Get the (true) wavelet coefficients and their variances.

```{r compute-wcs}
wc.sim     <- titable(x.sim)$difftable
wc.var.sim <- titable(sig.cb^2)$sumtable
wc.true    <- titable(mu.sp)$difftable
```

Get shrunken estimates of the wavelet coefficients.

```{r compute-shrunk-wcs}
wc.sim.shrunk <- vector("list",10)
wc.pres       <- vector("list",10)
for(j in 0:(log2(n) - 1)){
  wc.sim.shrunk[[j+1]] <-
    ash(wc.sim[j+2,],sqrt(wc.var.sim[j+2,]),prior = "nullbiased",
        pointmass = TRUE,mixsd = NULL,mixcompdist = "normal",
        gridmult = 2,df = NULL)$result
  wc.pres[[j+1]] <- 1/sqrt(wc.var.sim[j+2,])
}
```

Summarize results
-----------------

Plot the distribution of the observed wavelet coefficients at a
coarser scale (scale = 1, orange) and at a finer scale (scale = 7,
dark blue).

```{r plot-wc-histogram, fig.height=4, fig.width=6}
par(cex.axis = 1,cex.lab = 1.5)
hist(wc.sim[4,],breaks = 2,xlab = "observed wavelet coefficients",
     xlim = c(-25,25),ylim = c(0,600),col = "darkblue",xaxp = c(-25,25,10),
     yaxp = c(0,600,6),main = "")
hist(wc.sim[10,],breaks = 40,add = TRUE,col = "darkorange")
```

Observe that the scale = 7 WCs are much more concentrated near zero
because the signal is smoother at this finer scale.

Plot the observed wavelet coefficients (at scales 1 and 7 only)
vs. the "shrunken" wavelet coefficients estimated by adaptive
shrinkage.

```{r wc-scatterplot, fig.height=5, fig.width=5}
par(cex.axis = 1,cex.lab = 1.5)
plot(c(),c(),xlab = "observed wavelet coefficients",
     ylab = "shrunken wavelet coefficients",
     xlim = c(-2.5,2.5),ylim = c(-2.5,2.5))
abline(0,1,lty = 1,col = "gray",lwd = 1)
points(wc.sim[10,],wc.sim.shrunk[[9]]$PosteriorMean,pch = 20,cex = 1,
       col = "darkorange")
points(wc.sim[4,],wc.sim.shrunk[[3]]$PosteriorMean,pch = 20,cex = 1,
       col = "darkblue")
```

Plot the observed wavelet coefficients (at scale 7 only) vs. the
"shrunken" wavelet coefficients estimated by adaptive shrinkage, and
how how the amount of shrinkage depends on the standard error (s.e.)
in the observations.

```{r}
wc.sig.3 <- 1/wc.pres[[3]]
p <- ggplot(data.frame(observed = wc.sim[4,],
                       shrunken = wc.sim.shrunk[[3]]$PosteriorMean,
                       se       = wc.sig.3),
            aes(x = observed,y = shrunken,col = se)) +
  geom_point(na.rm = TRUE) +
  xlim(c(-1,1)) +
  ylim(c(-1,1)) +
  scale_color_gradientn(colors = c("deepskyblue","darkblue")) +
  theme_cowplot()
print(p)
```

The scale = 7 WCs are strongly shrunk toward zero, whereas the scale =
1 WCs are not shrunk nearly as much; SMASH infers from the data that
the scale = 7 WCs are heavily concentrated around zero, and
consequently SMASH shrinks them more strongly.

Plot the ground-truth signal (mean function) and the signals
recovered by the TI thresholding and smash methods.

```{r}
par(cex.axis = 0.8)
plot(mu.sp,type = "l",col = "black",lwd = 2,xlab = "position",ylab = "",
     ylim = c(-0.05,1),xaxp = c(0,1024,4),yaxp = c(0,1,4))
lines(mu.ti,col = "dodgerblue",lwd = 1.5)
lines(mu.smash,col = "orangered",lwd = 1.5)
cat("Demo is over.\n")
```