### 2018-08-24 analysis/shiny/shiny.Rmd
Got Shiny app running.

---
title: "Graphs"
runtime: shiny
output: html_document
---

NOTE: To launch the Shiny app, run `rmarkdown::run("shiny.Rmd")`

The following code generates an RShiny app which displays boxplots of
MISEs from the full simulation study comparing SMASH to other
wavelet-based methods for denoising Gaussian data. As mentioned in the
companion paper, the results for a variety of test functions (both
mean and variance functions) and 2 signal-to-noise ratios (SNRs), 1
and 3, are presented here.

Scenarios denote a combination of mean function, SNR and variance
function. Specifically, each scenario is named as
[mean_function].[SNR].[variance_function], where the abbreviations are
given by:

Mean function:  
Spikes - sp  
Bumps - bump  
Blocks - blk  
Angles - ang  
Doppler - dop  
Blip - blip  
Corner - cor

Variance function:  
Constant function - v1  
Triple exponential - v2  
Doppler - v3  
Bumps - v4  
Clipped Blocks - v5

On the other hand, we also ran a variety of methods in the simulation
study. The methods are abbreviated as follows:

EbayesThresh - ebayesthresh  
BAMS - bams  
BlockJS - blockjs 
Neighblock - neighblock  
SureShrink - sure  
Bayesian Posterior Mean - postmean  
TI thresholding - tithresh  
SMASH - smash

In addition, EbayeThresh, BAMS, BlockJS, Neighblock, Sureshrink and
Bayesian Posterior Mean all assume homoskedastic noise and use the
Symm8 basis by default. TI thresholding was run with both the
homoskedastic (tithresh.homo) using the Symm8 (.s8) basis, as well as
an estimated variance using RMAD (tithresh.rmad) and SMASH
(tithresh.smash), with both the Haar (.haar) and the Symm8 (.s8)
bases. In addition, TI thresholding was also run with the true
variance provided (tithresh.true) using both the Haar (.haar) and the
Symm8 (.s8) bases.

We ran SMASH with the homoskedatic (smash.homo) assumption using both
Haar (haar) and Symm8 (.s8). In addition, SMASH was also run with the
variance jointly estimated, using either the JASH procedure
(smash.jash) or the ASH procedure (smash.haar and smash.s8 for the two
different bases). Paralleling TI thresholding, we also ran SMASH with
the true variance provided (smash.true) using both the Haar (.haar)
and the Symm8 (.s8) bases.

```{r, echo=FALSE}
library(ggplot2)
library(dplyr)
load("../../dsc/results/res.RData")
scenario_names = as.character(unique(res$scenario))
method_names = as.character(unique(res$method))

inputPanel(

    # Create the checkboxes and select them all by default
    checkboxGroupInput("scen.subset", "Choose Scenarios", 
                        choices  = scenario_names,
                        selected = "sp.1.v2"),

    checkboxGroupInput("method.subset", "Choose Methods", 
                        choices  = method_names,
                        selected = method_names),

    selectInput("criteria", "Choose Criteria", 
                        choices  = "mise",
                        selected = "mise")
)

renderPlot({
  res.filter = filter(res,scenario %in% input$scen.subset &
    method %in% input$method.subset)
  print(input)
  res.filter$value = res.filter[[input$criteria]]
  ggplot(res.filter, aes(method, value, color = method)) +
    geom_boxplot() + facet_grid(.~scenario) + 
    theme(axis.ticks = element_blank(), axis.text.x = element_blank())
})
```