### 2018-09-28 analysis/gaussian.mean.est.Rmd
I have a complete first draft of the gaussian.mean.est.Rmd analysis.

---
title: "Gaussian mean estimation in simulated data sets"
author: "Zhengrong Xing, Peter Carbonetto and Matthew Stephens"
output: workflowr::wflow_html
---

*TO DO: Give overview of this report.*

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,fig.align = "center")
```

## Set up environment

Load the ggplot2 and cowplot packages, and the functions definining
the mean and variances used to simulate the data.

```{r load-pkgs, message=FALSE, warning=FALSE}
library(ggplot2)
library(cowplot)
source("../code/signals.R")
```

## Load results

Load the results of the simulation experiments.

```{r load-results}
load("../output/gaus-dscr.RData")
```

## Simulated data with constant variances

This plot reproduces Fig. 2 of the manuscript comparing the
accuracy of estimated mean curves in the data sets simulated from
the "Spikes" mean function with constant variance.

First, extract the results used to generate this plot.

```{r plot-1-extract-data}
homo.data.smash <-
  res[res$.id    == "sp.3.v1" &
      res$method == "smash.s8",]
homo.data.smash.homo <-
  res[res$.id    == "sp.3.v1" &
      res$method == "smash.homo.s8",]
homo.data.tithresh <-
  res[res$.id == "sp.3.v1" &
      res$method == "tithresh.homo.s8",]
homo.data.ebayes <-
  res[res$.id    == "sp.3.v1" &
      res$method == "ebayesthresh",]
homo.data.smash.true <-
  res[res$.id == "sp.3.v1" &
  res$method  == "smash.true.s8",]
homo.data <-
  res[res$.id == "sp.3.v1" &
  (res$method == "smash.s8" |
   res$method == "ebayesthresh" |
   res$method == "tithresh.homo.s8"),]
```

Transform these results into a data frame suitable for ggplot2.

```{r plot-1-transform-data}
pdat <-
  rbind(data.frame(method      = "smash",
                   method.type = "est",
                   mise        = homo.data.smash$mise),
        data.frame(method      = "smash.homo",
                   method.type = "homo",
                   mise        = homo.data.smash.homo$mise),
        data.frame(method      = "tithresh",
                   method.type = "homo",
                   mise        = homo.data.tithresh$mise),
        data.frame(method      = "ebayesthresh",
                   method.type = "homo",
                   mise        = homo.data.ebayes$mise),
        data.frame(method      = "smash.true",
                   method.type = "true",
                   mise        = homo.data.smash.true$mise))
pdat <-
  transform(pdat,
            method = factor(method,
                            names(sort(tapply(pdat$mise,pdat$method,mean),
                                       decreasing = TRUE))))
```

Create the combined boxplot and violin plot using ggplot2.

```{r plot-1-create, fig.height=4, fig.width=5}
p <- ggplot(pdat,aes(x = method,y = mise,fill = method.type)) +
     geom_violin(fill = "skyblue",color = "skyblue") +
     geom_boxplot(width = 0.15,outlier.shape = NA) +
     scale_y_continuous(breaks = seq(6,16,2)) +
     scale_fill_manual(values = c("darkorange","dodgerblue","gold"),
                       guide = FALSE) +
     coord_flip() +
     labs(x = "",y = "MISE") +
     theme(axis.line = element_blank(),
           axis.ticks.y = element_blank())
print(p)
```

From this plot, we see that three versions of SMASH outperformed
EbayesThresh and TI thresholding.

Next, we compare the same methods in simulated data sets with
heteroskedastic errors.

## Simulated data with heteroskedastic errors: "Spikes" mean signal and "Clipped Blocks" variance

In this scenario, data sets are simulated using the "Spikes" mean
function and the "Clipped Blocks" variance function. This plot shows
the mean function as a block line, and the +/- 2 standard deviations
as orange lines:

```{r spikes-signal, fig.height=4, fig.width=6}
t         <- (1:1024)/1024
mu        <- spikes.fn(t,"mean")
sigma.ini <- sqrt(cblocks.fn(t,"var"))
sd.fn     <- sigma.ini/mean(sigma.ini) * sd(mu)/3
par(cex.axis = 1,cex.lab = 1.25)
plot(mu,type = "l", ylim = c(-0.05,1),xlab = "position",ylab = "",
     lwd = 1.75,xaxp = c(0,1024,4),yaxp = c(0,1,4))
lines(mu + 2*sd.fn,col = "darkorange",lty = 5,lwd = 1.75)
lines(mu - 2*sd.fn,col = "darkorange",lty = 5,lwd = 1.75)
```

Extract the results from running these simulations.

```{r plot-2-extract-data}
hetero.data.smash <-
  res[res$.id == "sp.3.v5" & res$method == "smash.s8",]
hetero.data.smash.homo <-
  res[res$.id == "sp.3.v5" & res$method == "smash.homo.s8",]
hetero.data.tithresh.homo <-
  res[res$.id == "sp.3.v5" & res$method == "tithresh.homo.s8",]
hetero.data.tithresh.rmad <-
  res[res$.id == "sp.3.v5" & res$method == "tithresh.rmad.s8",]
hetero.data.tithresh.smash <-
  res[res$.id == "sp.3.v5" & res$method == "tithresh.smash.s8",]
hetero.data.tithresh.true <-
  res[res$.id == "sp.3.v5" & res$method == "tithresh.true.s8",]
hetero.data.ebayes <-
  res[res$.id == "sp.3.v5" & res$method == "ebayesthresh",]
hetero.data.smash.true <-
  res[res$.id == "sp.3.v5" & res$method == "smash.true.s8",]
```

Transform these results into a data frame suitable for ggplot2.

```{r plot-2-transform-data}
pdat <-
  rbind(data.frame(method      = "smash",
                   method.type = "est",
                   mise        = hetero.data.smash$mise),
        data.frame(method      = "smash.homo",
                   method.type = "homo",
                   mise        = hetero.data.smash.homo$mise),
        data.frame(method      = "tithresh.rmad",
                   method.type = "tithresh",
                   mise        = hetero.data.tithresh.rmad$mise),
        data.frame(method      = "tithresh.smash",
                   method.type = "tithresh",
                   mise        = hetero.data.tithresh.smash$mise),
        data.frame(method      = "tithresh.true",
                   method.type = "tithresh",
                   mise        = hetero.data.tithresh.true$mise),
        data.frame(method      = "ebayesthresh",
                   method.type = "homo",
                   mise        = hetero.data.ebayes$mise),
        data.frame(method      = "smash.true",
                   method.type = "true",
                   mise        = hetero.data.smash.true$mise))
pdat <-
  transform(pdat,
            method = factor(method,
                            names(sort(tapply(pdat$mise,pdat$method,mean),
                                       decreasing = TRUE))))
```

Create the combined boxplot and violin plot using ggplot2.

```{r plot-2-create, fig.height=4, fig.width=5}
p <- ggplot(pdat,aes(x = method,y = mise,fill = method.type)) +
     geom_violin(fill = "skyblue",color = "skyblue") +
     geom_boxplot(width = 0.15,outlier.shape = NA) +
     scale_fill_manual(values=c("darkorange","dodgerblue","limegreen","gold"),
                       guide = FALSE) +
     coord_flip() +
     scale_y_continuous(breaks = seq(10,70,10)) +
     labs(x = "",y = "MISE") +
     theme(axis.line = element_blank(),
           axis.ticks.y = element_blank())
print(p)
```

In this scenario, we see that SMASH, allowing for heteroskedastic
errors, outperforms EbayesThresh and all variants of TI thresholding
(including TI thresholding with the true variance). Further, SMASH
performs almost as well when estimating the variance compared to when
provided with the true variance.

## Simulated data with heteroskedastic errors: "Corner" mean signal and "Doppler" variance

In this next scenario, data sets are simulated using the "Corner" mean
function and the "Doppler" variance function. This plot shows the mean
function as a block line, and the +/- 2 standard deviations as orange
lines:

```{r corner-signal, fig.height=4, fig.width=6}
mu        <- cor.fn(t,"mean") 
sigma.ini <- sqrt(doppler.fn(t,"var"))
sd.fn     <- sigma.ini/mean(sigma.ini) * sd(mu)/3
plot(mu,type = "l", ylim = c(-0.05,1),xlab = "position",ylab = "",
     lwd = 1.75,xaxp = c(0,1024,4),yaxp = c(0,1,4))
lines(mu + 2*sd.fn,col = "darkorange",lty = 5,lwd = 1.75)
lines(mu - 2*sd.fn,col = "darkorange",lty = 5,lwd = 1.75)
```

Extract the results from running these simulations.

```{r plot-3-extract-data}
hetero.data.smash.2 <-
  res[res$.id == "cor.3.v3" & res$method == "smash.s8",]
hetero.data.smash.homo.2 <-
  res[res$.id == "cor.3.v3" & res$method == "smash.homo.s8",]
hetero.data.tithresh.homo.2 <-
  res[res$.id == "cor.3.v3" & res$method == "tithresh.homo.s8",]
hetero.data.tithresh.rmad.2 <-
  res[res$.id == "cor.3.v3" & res$method == "tithresh.rmad.s8",]
hetero.data.tithresh.smash.2 <-
  res[res$.id == "cor.3.v3" & res$method == "tithresh.smash.s8",]
hetero.data.tithresh.true.2 <-
  res[res$.id == "cor.3.v3" & res$method == "tithresh.true.s8",]
hetero.data.ebayes.2 <-
  res[res$.id == "cor.3.v3" & res$method == "ebayesthresh",]
hetero.data.smash.true.2 <-
  res[res$.id == "cor.3.v3" & res$method == "smash.true.s8",]
```

Transform these results into a data frame suitable for ggplot2.

```{r plot-3-transform-data}
pdat <-
  rbind(data.frame(method      = "smash",
                   method.type = "est",
                   mise        = hetero.data.smash.2$mise),
        data.frame(method      = "smash.homo",
                   method.type = "homo",
                   mise        = hetero.data.smash.homo.2$mise),
        data.frame(method      = "tithresh.rmad",
                   method.type = "tithresh",
                   mise        = hetero.data.tithresh.rmad.2$mise),
        data.frame(method      = "tithresh.smash",
                   method.type = "tithresh",
                   mise        = hetero.data.tithresh.smash.2$mise),
        data.frame(method      = "tithresh.true",
                   method.type = "tithresh",
                   mise        = hetero.data.tithresh.true.2$mise),
        data.frame(method      = "ebayesthresh",
                   method.type = "homo",
                   mise        = hetero.data.ebayes.2$mise),
        data.frame(method      = "smash.true",
                   method.type = "true",
                   mise        = hetero.data.smash.true.2$mise))
pdat <-
  transform(pdat,
            method = factor(method,
                            names(sort(tapply(pdat$mise,pdat$method,mean),
                                       decreasing = TRUE))))
```

Create the combined boxplot and violin plot using ggplot2.

```{r plot-3-create, fig.height=4, fig.width=5}
p <- ggplot(pdat,aes(x = method,y = mise,fill = method.type)) +
     geom_violin(fill = "skyblue",color = "skyblue") +
     geom_boxplot(width = 0.15,outlier.shape = NA) +
     scale_fill_manual(values=c("darkorange","dodgerblue","limegreen","gold"),
                       guide = FALSE) +
     coord_flip() +
     scale_y_continuous(breaks = seq(1,5)) +
     labs(x = "",y = "MISE") +
     theme(axis.line = element_blank(),
           axis.ticks.y = element_blank())
print(p)
```

Similar to the "spikes" data sets, we see that the SMASH method
allowing for heteroskedastic variances outperforms the TI thresholding
and EbayesThresh approaches.