### 2018-10-10 analysis/motorcycle.Rmd
wflow_publish("motorcycle.Rmd")

---
title: "Motorcycle Acceleration example"
author: "Zhengrong Xing, Peter Carbonetto and Matthew Stephens"
output: workflowr::wflow_html
---

An illustration of "smoothing via adaptive shrinkage" (SMASH) applied to the
Motorcycle Accident data. This implements the "illustrative application"
presented in Sec. 5.1 of the manuscript.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,fig.align = "center")
```

Set up environment
------------------

Load the MASS, scales, wavethresh, EbayesThresh and smashr
packages. The MASS package is loaded only for the Motorcycle
Accident data. Some additional functions are defined in
`functions.motorcycle.R`.

```{r load-pkgs, message=FALSE}
library(MASS)
library(lattice)
library(smashr)
library(scales)
library(wavethresh)
library(EbayesThresh)
source("../code/motorcycle.functions.R")
```

Prepare data for SMASH
----------------------

Load the motorcycle data from the MASS package, and order the data
points by time.

```{r prepare-data}
data(mcycle)
x.ini.mc <- sort(mcycle$times)
y.ini.mc <- mcycle$accel[order(mcycle$times)]
```

Run SMASH
---------

Apply SMASH to the Motorcycle Accident data set.

```{r run-smash}
res.mc <- smash.wrapper(x.ini.mc,y.ini.mc)
```

Summarize results of SMASH analysis
-----------------------------------

Create a plot showing the Motorcycle Accident data and the smash
estimates (with the dashed red lines showing the confidence
intervals).

```{r plot-smash-estimates, fig.height=5, fig.width=7}
plot(res.mc$x,res.mc$mu.est,type = "l",
     ylim = c(min(res.mc$y - 2 * sqrt(res.mc$var.est)),
              max(res.mc$y + 2 * sqrt(res.mc$var.est))),
     xlab = "time (ms)", ylab = "acceleration (g)",lwd = 2,
	 col = "darkorange",xlim = c(0,60),xaxp = c(0,60,6))
lines(res.mc$x, res.mc$mu.est + 2*sqrt(res.mc$var.est),lty = 5,
	 lwd = 2,col = "dodgerblue")
lines(res.mc$x,res.mc$mu.est - 2*sqrt(res.mc$var.est),
      lty = 5,lwd = 2,col = "dodgerblue")
points(res.mc$x,res.mc$y,pch = 1,cex = 1,col = "black")
```

Apply SMASH again, this time assuming equal variances
-----------------------------------------------------

Apply SMASH (assuming with equal variances) to the Motorcycle Accident
data set.

```{r run-smash-homo}
res.cons.mc <- smash.cons.wrapper(x.ini.mc,y.ini.mc)
```

Apply TI thresholding to the Motorcycle Accident data
---------------------------------------------------

Apply TI thresholding to the Motorcycle Accident data set.

```{r run-tithresh}
res.ti.cons.mc <- tithresh.cons.wrapper(x.ini.mc,y.ini.mc)
```

Apply TI thresholding to the Motorcycle Accident data, this time using
the variances estimated by SMASH.

```{r run-tithresh-with-smash-var}
res.ti.mc <- tithresh.wrapper(x.ini.mc,y.ini.mc)
```

Compare results from all methods
--------------------------------

In this second plot, we compare the mean estimate provided by SMASH
(with heteroskedastic variances; orange line) against homoskedastic
SMASH (dotted magenta line), TI thresholding (dark blue line), and TI
thresholding when the variances have been estimated by SMASH (dotted
green line).

```{r plot-all-estimates, fig.height=5, fig.width=7}
plot(res.mc$x,res.mc$mu.est,type = "l",
     ylim = c(min(res.mc$y - 2 * sqrt(res.mc$var.est)),
              max(res.mc$y + 2 * sqrt(res.mc$var.est))),
	 xlab = "time (ms)",ylab = "acceleration (g)",lwd = 2,
	 col = "darkorange",xlim = c(0,60),xaxp = c(0,60,6))
lines(res.cons.mc$x,res.cons.mc$mu.est,lwd = 2,lty = "dotted",col = "magenta")
lines(res.ti.cons.mc$x,res.ti.cons.mc$mu.est,lwd = 2,lty = "solid",
      col = "darkblue")
lines(res.ti.mc$x,res.ti.mc$mu.est,lwd = 2,col = "limegreen",lty = "dotted")
points(res.mc$x,res.mc$y,pch = 1,cex = 0.8,col = "black")
```