### 2018-10-19 analysis/chipseq.Rmd
wflow_publish("chipseq.Rmd")

---
title: "ChIP-seq example"
author: "Zhengrong Xing, Peter Carbonetto and Matthew Stephens"
output: workflowr::wflow_html
---

An illustration of "smoothing via adaptive shrinkage" (SMASH) applied
to chromatin immunoprecipitation sequencing ("ChIP-seq") data. This
implements the SMASH analysis presented in Sec. 5.2 of the manuscript.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,fig.align = "center")
```

Set up environment
------------------

We begin by loading the smashr, ggplot2 and cowplot packages, as well
as some additional functions used to implement the analysis below.

```{r load-pkgs, warning=FALSE, message=FALSE}
source("../code/chipseq.functions.R")
library(smashr)
library(ggplot2)
library(cowplot)
```

Load the ChIP-seq data
----------------------

The ChIP-seq data are sequencing read counts for transcription
factor YY1 in cell line GM12878, restricted to 880,001--1,011,072 bp
on chromosome 1. These data were collected as part of the ENCODE
("Encyclopedia Of DNA Elements") project.

```{r load-data}
load("../data/reg_880000_1011072.RData")
bppos  <- 880001:1011072
counts <- M[1,] + M[,2]
```

Note that there are two replicates of the GM12878 cell line, so we
analyze the combined read counts from both replicates, stored in the
`counts` vectors.

Run SMASH
---------

Next, we apply SMASH to the read counts to estimate the mean and
variance of the underlying signal. It may take several minutes to
complete this step.

```{r run-smash}
res <- smash.poiss(counts,post.var = TRUE)
```

Plot the SMASH estimates
------------------------

To provide a "baseline" to compare against the SMASH estimates, we
retrieve the peaks identified in the same ChIP-seq data using the MACS
software.

```{r load-macs-peaks}
macs.file <- "../data/Gm1287peaks_chr1_sorted.txt"
peaks <- read.macs.peaks(macs.file,min(bppos),max(bppos))
```

This next plot shows the intensity function estimated by SMASH (orange
line). The read count data are depicted as light blue circles, in
which the area of each circle is scaled by the number of data points
that fall within each 1.6-kb "bin". (We show counts summarized within
bins because there are too many data points to plot them
individually.)

The peaks identified by the MACS software are shown as red
triangles. (Specifically, these are the *mean positions* of the
identified peak intervals; the peak intervals are short enough that it
is not useful to show both the start and end positions of these
intervals.)

```{r plot-smash-estimates, fig.height=3, fig.width=9}
create.chipseq.plot(bppos/1e6,counts,res$est,(peaks$start + peaks$end)/2e6,
                    nbreaks = 80) +
  scale_x_continuous(limits = c(0.88,1.02),breaks = seq(0.88,1.02,0.02)) +
  scale_y_continuous(limits = c(-1,9),breaks = seq(0,8,2))
```

Based on this plot, it is clear that the the strongest SMASH intensity
estimates align very closely with the peaks found by MACS.

Intriguingly, the SMASH estimates also suggest the presence of several
additional weaker peaks that were not identified by MACS.