### 2018-11-06 analysis/spikesdemo.Rmd
Working on improvements to the Spikes demo; mainly in the setup instructions.

---
title: "Spikes demo"
author: "Zhengrong Xing, Peter Carbonetto and Matthew Stephens"
output: workflowr::wflow_html
---

This small demonstration illustrates key features of the SMASH method
on a small, simulated data set with spatially structured variances. This
code implements the example presented in Sec. 4.1 of the paper.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,fig.align = "center")
```

Initial setup instructions
--------------------------

To run this example on your own computer, please follow these setup
instructions. These instructions assume you already have R and/or
RStudio installed on your computer.

Download or clone the [git repository][smash-github] on your computer.

Launch R, and change the working directory to be the `analysis` folder
inside your local copy of the `smash-paper` git repository.

Install some of the packages used here, and in the code below:

```{r install-pkgs-1, eval=FALSE}
install.packages(c("devtools","ggplot2","cowplot"))
```

Finally, install the ashr and smashr packages from GitHub:

```{r install-pkgs-2, eval=FALSE}
library(devtools)
install_github("stephens999/ashr",upgrade_dependencies = FALSE,
               build_vignettes = TRUE)
install_github("stephenslab/smashr",upgrade_dependencies = FALSE,
               build_vignettes = TRUE)
```

Set up R environment
--------------------

We begin by loading the ashr, smashr, scales, ggplot2 and cowplot
packages, as well as some additional functions used in the code below.

```{r load-pkgs, warning=FALSE, message=FALSE}
library(ashr)
library(smashr)
library(scales)
library(ggplot2)
library(cowplot)
source("../code/misc.R")
```

Define the "spikes" mean function
---------------------------------

Variable `mu.sp` defines the mean signal.

```{r define-spikes}
n <- 1024
t <- 1:n/n
spike.f <-
  function (x) (0.75 * exp(-500   * (x - 0.23)^2) +
                1.5  * exp(-2000  * (x - 0.33)^2) +
                3    * exp(-8000  * (x - 0.47)^2) +
                2.25 * exp(-16000 * (x - 0.69)^2) +
                0.5  * exp(-32000 * (x - 0.83)^2))
mu.sp <- spike.f(t)
mu.sp <- (1 + mu.sp)/5
```

Simulate data
-------------

Simulate the noisy observations from the normal distribution with
means that are given by the "spikes" function, and with
spatially-structured variances given by the "clipped blocks" function.

```{r simulate-data}
pos <- c(0.1,0.13,0.15,0.23,0.25,0.4,0.44,0.65,0.76,0.78,0.81)
hgt <- 2.88/5 * c(4,-5,3,-4,5,-4.2,2.1,4.3,-3.1,2.1,-4.2)
sig.cb <- rep(0,length(t))
for (j in 1:length(pos)) 
  sig.cb <- sig.cb + (1 + sign(t - pos[j])) * (hgt[j]/2)
sig.cb[sig.cb < 0] <- 0
sig.cb <- 0.1 + (sig.cb - min(sig.cb))/max(sig.cb)
rsnr   <- sqrt(3)
sig.cb <- sig.cb/mean(sig.cb) * sd(mu.sp)/rsnr^2
x.sim  <- rnorm(n,mu.sp,sig.cb)
```

Plot the simulated data set.

```{r plot-simulated-data, fig.height=5, fig.width=10}
par(cex.axis = 1,cex.lab = 1.25)
plot(mu.sp,type = 'l',ylim = c(-0.05,1),xlab = "position",
     ylab = "",lwd = 1.7,xaxp = c(0,1024,4),yaxp = c(0,1,4))
lines(mu.sp + 2*sig.cb,col = "darkorange",lty = 5,lwd = 1.8)
lines(mu.sp - 2*sig.cb,col = "darkorange",lty = 5,lwd = 1.8)
points(x.sim,cex = 0.7,pch = 16,col = "darkblue")
```

This plot shows the "spikes" mean function (the black line), +/- 2
standard deviations (the orange lines), and the simulated data (the
blue points).

Run SMASH and TI thresholding
-----------------------------

Now we apply SMASH, as well as the translation invariant (TI)
thresholding method, to the "spikes" data. Here we run the TI
thresholding method twice---once when the variance is assumes to be
constant ("homoskedastic"), and once when it is estimated using the
RMAD algorithm. The first method yields very poor estimates, so they
are not shown in the plots below.

The SMASH method is able to estimate the variance.

```{r denoise-signals}
sig.est  <- sqrt(2/(3 * (n - 2)) *
              sum((1/2 * x.sim[1:(n-2)] - x.sim[2:(n-1)] + x.sim[3:n])^2/2))
mu.smash <- smash(x.sim,family = "DaubLeAsymm",filter.number = 8)
mu.ti.homo <- ti.thresh(x.sim,sigma = sig.est,family = "DaubLeAsymm",
                        filter.number = 8)
mu.ti    <- ti.thresh(x.sim,method = "rmad",family = "DaubLeAsymm",
                      filter.number = 8)
```

It should take at the very most a few minutes to run SMASH and TI
thresholding.

Get the (true) wavelet coefficients and their variances.

```{r compute-wcs}
wc.sim     <- titable(x.sim)$difftable
wc.var.sim <- titable(sig.cb^2)$sumtable
wc.true    <- titable(mu.sp)$difftable
```

Get the shrunken estimates of the wavelet coefficients.

```{r compute-shrunk-wcs}
wc.sim.shrunk <- vector("list",10)
wc.pres       <- vector("list",10)
for(j in 0:(log2(n) - 1)){
  wc.sim.shrunk[[j+1]] <-
    ash(wc.sim[j+2,],sqrt(wc.var.sim[j+2,]),prior = "nullbiased",
        pointmass = TRUE,mixsd = NULL,mixcompdist = "normal",
        gridmult = 2,df = NULL)$result
  wc.pres[[j+1]] <- 1/sqrt(wc.var.sim[j+2,])
}
```

Summarize results
-----------------

Plot the distribution of the observed wavelet coefficients at a
coarser scale (scale = 1, orange) and at a finer scale (scale = 7,
dark blue).

```{r plot-wc-histogram, fig.height=4, fig.width=6}
par(cex.axis = 1,cex.lab = 1.25)
hist(wc.sim[4,],breaks = 2,xlab = "observed wavelet coefficients",
     xlim = c(-25,25),ylim = c(0,600),col = "darkblue",xaxp = c(-25,25,10),
     yaxp = c(0,600,6),main = "")
hist(wc.sim[10,],breaks = 40,add = TRUE,col = "darkorange")
```

Observe that the scale = 7 WCs (dark blue) are much more concentrated
near zero because the signal is smoother at this finer scale.

Plot the observed wavelet coefficients (at scales 1 and 7 only)
vs. the "shrunken" wavelet coefficients estimated by adaptive
shrinkage.

```{r wc-scatterplot, fig.height=5, fig.width=5}
par(cex.axis = 1,cex.lab = 1.25)
plot(c(),c(),xlab = "observed wavelet coefficients",
     ylab = "shrunken wavelet coefficients",
     xlim = c(-2.5,2.5),ylim = c(-2.5,2.5))
abline(0,1,lty = 1,col = "gray",lwd = 1)
points(wc.sim[10,],wc.sim.shrunk[[9]]$PosteriorMean,pch = 20,cex = 1,
       col = "darkorange")
points(wc.sim[4,],wc.sim.shrunk[[3]]$PosteriorMean,pch = 20,cex = 1,
       col = "darkblue")
```

Observe that the scale = 7 WCs (dark blue) are strongly shrunk toward
zero, whereas the scale = 1 WCs (orange) are not shrunk nearly as
much; SMASH infers from the data that the scale = 7 WCs are heavily
concentrated around zero, and consequently SMASH shrinks them more
strongly.

Plot the observed wavelet coefficients (at scale 7 only) vs. the
"shrunken" wavelet coefficients estimated by adaptive shrinkage in
order to see how the amount of shrinkage depends on the standard error
(s.e.) in the observations.

```{r wc-scatterplot-2, fig.height=3.75, fig.width=5}
wc.sig.3 <- 1/wc.pres[[3]]
p <- ggplot(data.frame(observed = wc.sim[4,],
                       shrunken = wc.sim.shrunk[[3]]$PosteriorMean,
                       se       = wc.sig.3),
            aes(x = observed,y = shrunken,col = se)) +
  geom_point(na.rm = TRUE) +
  xlim(c(-1,1)) +
  ylim(c(-1,1)) +
  scale_color_gradientn(colors = c("deepskyblue","darkblue")) +
  theme_cowplot()
print(p)
```

From this plot, it is evident that SMASH shrinks WCs differently
depending on their precision; specifically, WCs that are less
precise---i.e., higher standard error (s.e.)---are shrunk more
strongly toward zero.

Finally, we plot the ground-truth signal (the mean function, drawn as
a black line) and the signals recovered by TI thresholding (light blue
line) and SMASH (the red line).

```{r plot-mean-signal-vs-estimates, fig.height=5, fig.width=10}
par(cex.axis = 1)
plot(mu.sp,type = "l",col = "black",lwd = 3,xlab = "position",ylab = "",
     ylim = c(-0.05,1),xaxp = c(0,1024,4),yaxp = c(0,1,4))
lines(mu.ti,col = "dodgerblue",lwd = 3)
lines(mu.smash,col = "orangered",lwd = 3)
```

In this example, the smoothed signal from SMASH is noticeably more
accurate than using TI thresholding (in which the variance is
estimated by running the median absolute deviation, or "RMAD",
method). The TI thresholding estimate shows notable artifacts.

[smash-github]: https://github.com/stephenslab/smash-paper