### 2018-11-08 analysis/gaussvarest.Rmd
Moved around some files and revised TOC in home page.

# ---
# title: "Gaussian variance estimation in simulated data sets"
# ---
#
# This script implements the "Gaussian variance estimation" simulation
# experiments in the paper. In particular, we compare the Mean Field
# Variational Bayes method (MFVB) against SMASH in two scenarios. See
# the paper for the full description of these scenarios.
#
# The figure and table generated at the end of this script should
# match up with the figure and table shown in the paper.
#
# Note that this script could take an hour or several hours to
# complete as it runs the two methods on 100 simulated data sets for
# each of the two scenarios.
#
# We thank Menictas & Wand for generously sharing code that was used
# to implement these experiments.
#

# -----
# TO DO: Remove steps that are not included in the paper.
# -----

# SET UP ENVIRONMENT
# ------------------
# Load the smashr package, and some functions used in the analysis.
library(smashr)
source("../code/mfvb.functions.R")

# SCRIPT PARAMETERS
# -----------------
# Number of data sets simulated in the first (nsim1) and second
# (nsim2) simulation scenarios.
nsim1 <- 100
nsim2 <- 100

# Set hyperparameters.
Au.hyp      <- 1e5
Av.hyp      <- 1e5
sigsq.gamma <- 1e10
sigsq.beta  <- 1e10

# Set colours.
mainCol <- "darkslateblue"
ptCol   <- "paleturquoise3"
lineCol <- "skyblue"
axisCol <- "black"

# Set plotting size parameters.
cex.pt      <- 0.75
cex.mainVal <- 1.7
cex.labVal  <- 1.3
xlabVal     <- "x"

# FIRST SIMULATION SCENARIO
# -------------------------
mse.mu.uneven.mfvb <- 0
mse.mu.uneven.haar <- 0
mse.mu.uneven.s8   <- 0

mse.sd.uneven.mfvb <- 0
mse.sd.uneven.haar <- 0
mse.sd.uneven.s8   <- 0
mse.sd.uneven.j    <- 0
mse.sd.uneven.s8.j <- 0

lmse.sd.uneven.mfvb <- 0
lmse.sd.uneven.haar <- 0
lmse.sd.uneven.s8   <- 0
lmse.sd.uneven.j    <- 0
lmse.sd.uneven.s8.j <- 0

# Repeat for each data set simulated in the first simulation scenario.
cat(sprintf("Running %d simulations for Simulation Scenario 1.\n",nsim1))
for (j in 1:nsim1) {
  cat(sprintf("%d ",j))

  # Create the simulated data set
  # -----------------------------
  set.seed(3*j)
  n     <- 500
  xOrig <- runif(n)
  set.seed(3*j)
  yOrig <- fTrue(xOrig) + sqrt(exp(loggTrue(xOrig)))*rnorm(n)

  aOrig  <- min(xOrig)
  bOrig  <- max(xOrig)
  mean.x <- mean(xOrig)
  sd.x   <- sd(xOrig)
  mean.y <- mean(yOrig)
  sd.y   <- sd(yOrig)

  a <- (aOrig - mean.x)/sd.x
  b <- (bOrig - mean.x)/sd.x
  x <- (xOrig - mean.x)/sd.x
  y <- (yOrig - mean.y)/sd.y

  numIntKnotsU <- 17
  intKnotsU <- quantile(x,seq(0,1,length=numIntKnotsU+2)[-c(1,numIntKnotsU+2)])
  Zu        <- ZOSull(x,intKnots=intKnotsU,range.x=c(a,b))
  numKnotsU <- ncol(Zu)

  numIntKnotsV <- numIntKnotsU 
  intKnotsV <-
    quantile(x,seq(0,1,length = numIntKnotsV + 2)[-c(1,numIntKnotsV+2)])
  Zv        <- ZOSull(x,intKnots=intKnotsV,range.x=c(a,b))
  numKnotsV <- ncol(Zv)

  # Run Mean Field Variational Bayes (MFVB)
  # ---------------------------------------
  X     <- cbind(rep(1,n),x)
  Cumat <- cbind(X,Zu)
  Cvmat <- cbind(X,Zv)
  ncX   <- ncol(X)
  ncZu  <- ncol(Zu)
  ncZv  <- ncol(Zv)
  ncCu  <- ncol(Cumat)
  ncCv  <- ncol(Cvmat)

  MFVBfit <- meanVarMFVB(y,X,ncZu,ncZv,Au.hyp,Av.hyp,
                         sigsq.gamma,sigsq.beta)

  ng     <- 201
  xgOrig <- seq(aOrig,bOrig,length=ng)
  xg     <- (xgOrig - mean.x)/sd.x
  Xg     <- cbind(rep(1,ng),xg)
  Zug    <- ZOSull(xg,intKnots=intKnotsU,range.x=c(a,b))
  Cug    <- cbind(Xg,Zug)
  Zvg    <- ZOSull(xg,intKnots=intKnotsV,range.x=c(a,b))
  Cvg    <- cbind(Xg,Zvg)

  mu.q.nu       <- MFVBfit$mu.q.nu
  mu.q.omega    <- MFVBfit$mu.q.omega
  Sigma.q.nu    <- MFVBfit$Sigma.q.nu
  Sigma.q.omega <- MFVBfit$Sigma.q.omega

  fhatMFVBg        <- Cug%*%mu.q.nu
  fhatMFVBgOrig    <- fhatMFVBg*sd.y + mean.y
  logghatMFVBg     <- Cvg%*%mu.q.omega 
  logghatMFVBgOrig <- logghatMFVBg + 2*log(sd.y)

  sdloggMFVBgOrig      <- sqrt(diag(Cvg%*%Sigma.q.omega%*%t(Cvg))) 
  credLowloggMFVBgOrig <- logghatMFVBgOrig - qnorm(0.975)*sdloggMFVBgOrig
  credUpploggMFVBgOrig <- logghatMFVBgOrig + qnorm(0.975)*sdloggMFVBgOrig

  sqrtghatMFVBg     <- exp(0.5*Cvg %*% mu.q.omega 
                           + 0.125*diag(Cvg%*%Sigma.q.omega%*%t(Cvg)))
  sqrtghatMFVBgOrig <- sqrtghatMFVBg*sd.y

  # Run SMASH
  # ---------
  x.mod <- unique(sort(xOrig))
  y.mod <- 0
  for(i in 1:length(x.mod))
    y.mod[i] <- median(yOrig[xOrig == x.mod[i]])

  y.exp   <- c(y.mod,y.mod[length(y.mod):(2*length(y.mod)-2^9+1)])
  y.final <- c(y.exp,y.exp[length(y.exp):1])

  mu.est  <- smash.gaus(y.final,filter.number=1,family="DaubExPhase")
  var.est <- smash.gaus(y.final,v.est=TRUE)
  mu.est  <- mu.est[1:500]
  var.est <- var.est[1:500]

  mu.est.inter  <- approx(x.mod,mu.est,xgOrig,'linear')$y
  var.est.inter <- approx(x.mod,var.est,xgOrig,'linear')$y

  mse.mu.uneven.mfvb[j] <- mean((fhatMFVBgOrig - fTrue(xgOrig))^2)
  mse.mu.uneven.haar[j] <- mean((mu.est.inter - fTrue(xgOrig))^2)

  mse.sd.uneven.mfvb[j]=mean((sqrtghatMFVBgOrig-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.uneven.haar[j]=mean((sqrt(var.est.inter)-exp((loggTrue(xgOrig))/2))^2)
  lmse.sd.uneven.mfvb[j]=mean((log(sqrtghatMFVBgOrig)-(loggTrue(xgOrig))/2)^2)
  lmse.sd.uneven.haar[j]=
    mean((log(sqrt(var.est.inter))-(loggTrue(xgOrig))/2)^2)

  mu.est       <- smash.gaus(y.final,filter.number=8,family="DaubLeAsymm")
  var.est      <- smash.gaus(y.final,v.est=TRUE,v.basis=TRUE,filter.number=8,
                             family="DaubLeAsymm")
  var.est.s8.j <- smash.gaus(y.final,v.est=TRUE,v.basis=TRUE,jash=TRUE,
                             filter.number=8,family="DaubLeAsymm")
  var.est.j    <- smash.gaus(y.final,v.est=TRUE,jash=TRUE)

  mu.est       <- mu.est[1:500]
  var.est      <- var.est[1:500]
  var.est.j    <- var.est.j[1:500]
  var.est.s8.j <- var.est.s8.j[1:500]

  mu.est.inter       <- approx(x.mod,mu.est,xgOrig,'linear')$y
  var.est.inter      <- approx(x.mod,var.est,xgOrig,'linear')$y
  var.est.inter.j    <- approx(x.mod,var.est.j,xgOrig,'linear')$y
  var.est.inter.s8.j <- approx(x.mod,var.est.s8.j,xgOrig,'linear')$y

  mse.mu.uneven.s8[j]=mean((mu.est.inter-fTrue(xgOrig))^2)
  mse.sd.uneven.s8[j]=mean((sqrt(var.est.inter)-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.uneven.j[j] =mean((sqrt(var.est.inter.j)-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.uneven.s8.j[j] =
    mean((sqrt(var.est.inter.s8.j) - exp((loggTrue(xgOrig))/2))^2)
  lmse.sd.uneven.s8[j]=mean((log(sqrt(var.est.inter))-(loggTrue(xgOrig))/2)^2)
  lmse.sd.uneven.j[j]=mean((log(sqrt(var.est.inter.j))-(loggTrue(xgOrig))/2)^2)
  lmse.sd.uneven.s8.j[j] =
    mean((log(sqrt(var.est.inter.s8.j))-(loggTrue(xgOrig))/2)^2)
}
cat("\n")

# SECOND SIMULATION SCENARIO
# --------------------------
mse.mu.even.mfvb <- 0
mse.mu.even.haar <- 0
mse.mu.even.s8   <- 0

mse.sd.even.mfvb <- 0
mse.sd.even.haar <- 0
mse.sd.even.s8   <- 0
mse.sd.even.j    <- 0
mse.sd.even.s8.j <- 0

lmse.sd.even.mfvb <- 0
lmse.sd.even.haar <- 0
lmse.sd.even.s8   <- 0
lmse.sd.even.j    <- 0
lmse.sd.even.s8.j <- 0

# Repeat for each data set simulated in the second simulation scenario.
cat(sprintf("Running %d simulations for Simulation Scenario 2.\n",nsim2))
for (j in 1:nsim2) {
  cat(sprintf("%d ",j))

  # Create the simulated data set
  # -----------------------------
  n      <- 2^10
  xOrig  <- (1:n)/n
  set.seed(30*j)
  yOrig  <- fTrue(xOrig) + sqrt(exp(loggTrue(xOrig)))*rnorm(n)
  aOrig  <- min(xOrig)
  bOrig  <- max(xOrig)
  mean.x <- mean(xOrig)
  sd.x   <- sd(xOrig)
  mean.y <- mean(yOrig)
  sd.y   <- sd(yOrig)

  a <- (aOrig - mean.x)/sd.x
  b <- (bOrig - mean.x)/sd.x
  x <- (xOrig - mean.x)/sd.x
  y <- (yOrig - mean.y)/sd.y

  numIntKnotsU <- 17
  intKnotsU <- quantile(x,seq(0,1,length=numIntKnotsU+2)[-c(1,numIntKnotsU+2)])
  Zu        <- ZOSull(x,intKnots=intKnotsU,range.x=c(a,b))
  numKnotsU <- ncol(Zu)

  numIntKnotsV <- numIntKnotsU 
  intKnotsV <- quantile(x,seq(0,1,length=numIntKnotsV+2)[-c(1,numIntKnotsV+2)])
  Zv        <- ZOSull(x,intKnots=intKnotsV,range.x=c(a,b))
  numKnotsV <- ncol(Zv)

  # Run Mean Field Variational Bayes (MFVB)
  # ---------------------------------------
  X     <- cbind(rep(1,n),x)
  Cumat <- cbind(X,Zu)
  Cvmat <- cbind(X,Zv)
  ncX   <- ncol(X)
  ncZu  <- ncol(Zu)
  ncZv  <- ncol(Zv)
  ncCu  <- ncol(Cumat)
  ncCv  <- ncol(Cvmat)

  MFVBfit <- meanVarMFVB(y,X,ncZu,ncZv,Au.hyp,Av.hyp,
                       sigsq.gamma,sigsq.beta)

  ng     <- 2^10
  xgOrig <- seq(aOrig,bOrig,length=ng)
  xg  <- (xgOrig - mean.x)/sd.x
  Xg  <- cbind(rep(1,ng),xg)
  Zug <- ZOSull(xg,intKnots=intKnotsU,range.x=c(a,b))
  Cug <- cbind(Xg,Zug)
  Zvg <- ZOSull(xg,intKnots=intKnotsV,range.x=c(a,b))
  Cvg <- cbind(Xg,Zvg)

  mu.q.nu       <- MFVBfit$mu.q.nu
  mu.q.omega    <- MFVBfit$mu.q.omega
  Sigma.q.nu    <- MFVBfit$Sigma.q.nu
  Sigma.q.omega <- MFVBfit$Sigma.q.omega

  # Get the mean function estimate.
  fhatMFVBg     <- Cug %*% mu.q.nu
  fhatMFVBgOrig <- fhatMFVBg*sd.y + mean.y

  logghatMFVBg     <- Cvg%*%mu.q.omega 
  logghatMFVBgOrig <- logghatMFVBg + 2*log(sd.y)

  sdloggMFVBgOrig      <- sqrt(diag(Cvg%*%Sigma.q.omega%*%t(Cvg))) 
  credLowloggMFVBgOrig <- logghatMFVBgOrig - qnorm(0.975)*sdloggMFVBgOrig
  credUpploggMFVBgOrig <- logghatMFVBgOrig + qnorm(0.975)*sdloggMFVBgOrig

  sqrtghatMFVBg     <- exp(0.5*Cvg%*%mu.q.omega 
                           + 0.125*diag(Cvg%*%Sigma.q.omega%*%t(Cvg)))
  sqrtghatMFVBgOrig <- sqrtghatMFVBg*sd.y

  # Run SMASH
  # ---------
  mu.est  <- smash.gaus(yOrig,filter.number=1,family="DaubExPhase")
  var.est <- smash.gaus(yOrig,v.est=TRUE)

  mse.mu.even.mfvb[j] <- mean((fhatMFVBgOrig-fTrue(xgOrig))^2)
  mse.mu.even.haar[j] <- mean((mu.est-fTrue(xgOrig))^2)

  mse.sd.even.mfvb[j]  <- mean((sqrtghatMFVBgOrig-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.even.haar[j]  <- mean((sqrt(var.est)-exp((loggTrue(xgOrig))/2))^2)
  lmse.sd.even.mfvb[j] <- mean((log(sqrtghatMFVBgOrig)-(loggTrue(xgOrig))/2)^2)
  lmse.sd.even.haar[j] <- mean((log(sqrt(var.est))-(loggTrue(xgOrig))/2)^2)

  mu.est       <- smash.gaus(yOrig,filter.number=8,family="DaubLeAsymm")
  var.est      <- smash.gaus(yOrig,v.est=TRUE,v.basis=TRUE,filter.number=8,
                             family="DaubLeAsymm")
  var.est.s8.j <- smash.gaus(yOrig,v.est=TRUE,v.basis=TRUE,jash=TRUE,
                             filter.number=8,family="DaubLeAsymm")
  var.est.j    <- smash.gaus(yOrig,v.est=TRUE,jash=TRUE)

  mse.mu.even.s8[j]   <-mean((mu.est-fTrue(xgOrig))^2)
  mse.sd.even.s8[j]   <-mean((sqrt(var.est)-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.even.j[j]    <-mean((sqrt(var.est.j)-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.even.s8.j[j] <-mean((sqrt(var.est.s8.j)-exp((loggTrue(xgOrig))/2))^2)
  lmse.sd.even.s8[j]  <-mean((log(sqrt(var.est))-(loggTrue(xgOrig))/2)^2)
  lmse.sd.even.j[j]   <-mean((log(sqrt(var.est.j))-(loggTrue(xgOrig))/2)^2)
  lmse.sd.even.s8.j[j]<-mean((log(sqrt(var.est.s8.j))-(loggTrue(xgOrig))/2)^2)
}
cat("\n")

# PLOT MEAN & VARIANCE FUNCTIONS USED TO SIMULATE DATA
# ----------------------------------------------------
# Compare this plot against the one shown in Fig. 4 of the paper.
cat("Plotting mean and variance functions used to simulate data.\n")
xgrid <- (0:10000)/10000
plot(xgrid,fTrue(xgrid),type = "l",ylim = c(-5,5),ylab = "y",xlab = "X",
     lwd = 2)
lines(xgrid,fTrue(xgrid) + 2*sqrt(gTrue(xgrid)),col = "darkorange",lwd = 2)
lines(xgrid,fTrue(xgrid) - 2*sqrt(gTrue(xgrid)),col = "darkorange",lwd = 2)

# SUMMARIZE RESULTS
# -----------------
# Compare this table against Table 1 in the paper.
cat("Summarizing results of simulations.\n")
cat(sprintf("MSE averaged across %d simulations in Scenario 1,\n",nsim1))
cat(sprintf("and averaged across %d simulations in Scenario 2:\n",nsim2))
mse.table <- rbind(c(mean(mse.mu.uneven.mfvb),mean(mse.sd.uneven.mfvb),
                     mean(mse.mu.even.mfvb),mean(mse.sd.even.mfvb)),
                   c(mean(mse.mu.uneven.s8),mean(mse.sd.uneven.s8),
                     mean(mse.mu.even.s8),mean(mse.sd.even.s8)))
rownames(mse.table) <- c("MFVB","SMASH")
colnames(mse.table) <- c("mean","sd","mean","sd")
cat("         Scenario 1      Scenario 2   \n")
cat("      --------------- ----------------\n")
print(mse.table)