### 2018-11-09 analysis/gaussvarest.Rmd
wflow_publish("gaussvarest.Rmd")

---
title: "Gaussian variance estimation in simulated data sets"
author: "Zhengrong Xing, Peter Carbonetto and Matthew Stephens"
output: workflowr::wflow_html
---

This script implements the "Gaussian variance estimation" simulation
experiments in the paper. In particular, we compare the Mean Field
Variational Bayes method (MFVB) against SMASH in two scenarios.

The figure and table generated at the end of this script should
match up with the figure and table shown in the paper.

Running the code could take several hours to complete as it runs the
two methods on 100 simulated data sets for each of the two scenarios.

We thank M. Menictas & M. Wand for generously sharing code that was
used to implement these experiments.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,fig.align = "center")
```

Initial setup instructions
--------------------------

To run this example on your own computer, please follow these setup
instructions. These instructions assume you already have R and/or
RStudio installed on your computer.

First, download or clone the [git repository][smash-github] on your
computer.

Launch R, and change the working directory to be the "analysis" folder
inside your local copy of the git repository.

Finally, install the smashr package from GitHub:

```{r install-pkgs-2, eval=FALSE}
devtools::install_github("stephenslab/smashr",upgrade_dependencies = FALSE)
```

See the "Session Info" at the bottom for the versions of the software
and R packages that were used to generate the results shown below.

Set up R environment
--------------------

Load the smashr package, as well as some functions used in the
analysis below.

```{r load-pkgs}
library(smashr)
source("../code/mfvb.R")
```

Analysis settings
-----------------

Specify the number of data sets simulated in the first and second
simulation scenarios.

```{r analysis-settings-1}
nsim1 <- 10
nsim2 <- 10
```

Next, specify the hyperparameters used in running the MFVB method.

```{r analysis-settings-2}
Au.hyp      <- 1e5
Av.hyp      <- 1e5
sigsq.gamma <- 1e10
sigsq.beta  <- 1e10
```

These variables specify some colours used in the plots.

```{r analysis-settings-3}
mainCol <- "darkslateblue"
ptCol   <- "paleturquoise3"
lineCol <- "skyblue"
axisCol <- "black"
```

These are additional plotting parameters.

```{r analysis-settings-4}
cex.pt      <- 0.75
cex.mainVal <- 1.7
cex.labVal  <- 1.3
xlabVal     <- "x"
```

First simulation scenario
-------------------------

```{r simulation-1}
mse.mu.uneven.mfvb <- 0
mse.mu.uneven.haar <- 0
mse.mu.uneven.s8   <- 0

mse.sd.uneven.mfvb <- 0
mse.sd.uneven.haar <- 0
mse.sd.uneven.s8   <- 0
mse.sd.uneven.j    <- 0
mse.sd.uneven.s8.j <- 0

lmse.sd.uneven.mfvb <- 0
lmse.sd.uneven.haar <- 0
lmse.sd.uneven.s8   <- 0
lmse.sd.uneven.j    <- 0
lmse.sd.uneven.s8.j <- 0

# Repeat for each data set simulated in the first simulation scenario.
cat(sprintf("Running %d simulations for Simulation Scenario 1.\n",nsim1))
for (j in 1:nsim1) {
  cat(sprintf("%d ",j))

  # Create the simulated data set
  # -----------------------------
  set.seed(3*j)
  n     <- 500
  xOrig <- runif(n)
  set.seed(3*j)
  yOrig <- fTrue(xOrig) + sqrt(exp(loggTrue(xOrig)))*rnorm(n)

  aOrig  <- min(xOrig)
  bOrig  <- max(xOrig)
  mean.x <- mean(xOrig)
  sd.x   <- sd(xOrig)
  mean.y <- mean(yOrig)
  sd.y   <- sd(yOrig)

  a <- (aOrig - mean.x)/sd.x
  b <- (bOrig - mean.x)/sd.x
  x <- (xOrig - mean.x)/sd.x
  y <- (yOrig - mean.y)/sd.y

  numIntKnotsU <- 17
  intKnotsU <- quantile(x,seq(0,1,length=numIntKnotsU+2)[-c(1,numIntKnotsU+2)])
  Zu        <- ZOSull(x,intKnots=intKnotsU,range.x=c(a,b))
  numKnotsU <- ncol(Zu)

  numIntKnotsV <- numIntKnotsU 
  intKnotsV <-
    quantile(x,seq(0,1,length = numIntKnotsV + 2)[-c(1,numIntKnotsV+2)])
  Zv        <- ZOSull(x,intKnots=intKnotsV,range.x=c(a,b))
  numKnotsV <- ncol(Zv)

  # Run Mean Field Variational Bayes (MFVB)
  # ---------------------------------------
  X     <- cbind(rep(1,n),x)
  Cumat <- cbind(X,Zu)
  Cvmat <- cbind(X,Zv)
  ncX   <- ncol(X)
  ncZu  <- ncol(Zu)
  ncZv  <- ncol(Zv)
  ncCu  <- ncol(Cumat)
  ncCv  <- ncol(Cvmat)

  MFVBfit <- meanVarMFVB(y,X,ncZu,ncZv,Au.hyp,Av.hyp,
                         sigsq.gamma,sigsq.beta)

  ng     <- 201
  xgOrig <- seq(aOrig,bOrig,length=ng)
  xg     <- (xgOrig - mean.x)/sd.x
  Xg     <- cbind(rep(1,ng),xg)
  Zug    <- ZOSull(xg,intKnots=intKnotsU,range.x=c(a,b))
  Cug    <- cbind(Xg,Zug)
  Zvg    <- ZOSull(xg,intKnots=intKnotsV,range.x=c(a,b))
  Cvg    <- cbind(Xg,Zvg)

  mu.q.nu       <- MFVBfit$mu.q.nu
  mu.q.omega    <- MFVBfit$mu.q.omega
  Sigma.q.nu    <- MFVBfit$Sigma.q.nu
  Sigma.q.omega <- MFVBfit$Sigma.q.omega

  fhatMFVBg        <- Cug%*%mu.q.nu
  fhatMFVBgOrig    <- fhatMFVBg*sd.y + mean.y
  logghatMFVBg     <- Cvg%*%mu.q.omega 
  logghatMFVBgOrig <- logghatMFVBg + 2*log(sd.y)

  sdloggMFVBgOrig      <- sqrt(diag(Cvg%*%Sigma.q.omega%*%t(Cvg))) 
  credLowloggMFVBgOrig <- logghatMFVBgOrig - qnorm(0.975)*sdloggMFVBgOrig
  credUpploggMFVBgOrig <- logghatMFVBgOrig + qnorm(0.975)*sdloggMFVBgOrig

  sqrtghatMFVBg     <- exp(0.5*Cvg %*% mu.q.omega 
                           + 0.125*diag(Cvg%*%Sigma.q.omega%*%t(Cvg)))
  sqrtghatMFVBgOrig <- sqrtghatMFVBg*sd.y

  # Run SMASH
  # ---------
  x.mod <- unique(sort(xOrig))
  y.mod <- 0
  for(i in 1:length(x.mod))
    y.mod[i] <- median(yOrig[xOrig == x.mod[i]])

  y.exp   <- c(y.mod,y.mod[length(y.mod):(2*length(y.mod)-2^9+1)])
  y.final <- c(y.exp,y.exp[length(y.exp):1])

  mu.est  <- smash.gaus(y.final,filter.number=1,family="DaubExPhase")
  var.est <- smash.gaus(y.final,v.est=TRUE)
  mu.est  <- mu.est[1:500]
  var.est <- var.est[1:500]

  mu.est.inter  <- approx(x.mod,mu.est,xgOrig,'linear')$y
  var.est.inter <- approx(x.mod,var.est,xgOrig,'linear')$y

  mse.mu.uneven.mfvb[j] <- mean((fhatMFVBgOrig - fTrue(xgOrig))^2)
  mse.mu.uneven.haar[j] <- mean((mu.est.inter - fTrue(xgOrig))^2)

  mse.sd.uneven.mfvb[j]=mean((sqrtghatMFVBgOrig-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.uneven.haar[j]=mean((sqrt(var.est.inter)-exp((loggTrue(xgOrig))/2))^2)
  lmse.sd.uneven.mfvb[j]=mean((log(sqrtghatMFVBgOrig)-(loggTrue(xgOrig))/2)^2)
  lmse.sd.uneven.haar[j]=
    mean((log(sqrt(var.est.inter))-(loggTrue(xgOrig))/2)^2)

  mu.est       <- smash.gaus(y.final,filter.number=8,family="DaubLeAsymm")
  var.est      <- smash.gaus(y.final,v.est=TRUE,v.basis=TRUE,filter.number=8,
                             family="DaubLeAsymm")
  var.est.s8.j <- smash.gaus(y.final,v.est=TRUE,v.basis=TRUE,jash=TRUE,
                             filter.number=8,family="DaubLeAsymm")
  var.est.j    <- smash.gaus(y.final,v.est=TRUE,jash=TRUE)

  mu.est       <- mu.est[1:500]
  var.est      <- var.est[1:500]
  var.est.j    <- var.est.j[1:500]
  var.est.s8.j <- var.est.s8.j[1:500]

  mu.est.inter       <- approx(x.mod,mu.est,xgOrig,'linear')$y
  var.est.inter      <- approx(x.mod,var.est,xgOrig,'linear')$y
  var.est.inter.j    <- approx(x.mod,var.est.j,xgOrig,'linear')$y
  var.est.inter.s8.j <- approx(x.mod,var.est.s8.j,xgOrig,'linear')$y

  mse.mu.uneven.s8[j]=mean((mu.est.inter-fTrue(xgOrig))^2)
  mse.sd.uneven.s8[j]=mean((sqrt(var.est.inter)-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.uneven.j[j] =mean((sqrt(var.est.inter.j)-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.uneven.s8.j[j] =
    mean((sqrt(var.est.inter.s8.j) - exp((loggTrue(xgOrig))/2))^2)
  lmse.sd.uneven.s8[j]=mean((log(sqrt(var.est.inter))-(loggTrue(xgOrig))/2)^2)
  lmse.sd.uneven.j[j]=mean((log(sqrt(var.est.inter.j))-(loggTrue(xgOrig))/2)^2)
  lmse.sd.uneven.s8.j[j] =
    mean((log(sqrt(var.est.inter.s8.j))-(loggTrue(xgOrig))/2)^2)
}
cat("\n")
```

```
# SECOND SIMULATION SCENARIO
# --------------------------
mse.mu.even.mfvb <- 0
mse.mu.even.haar <- 0
mse.mu.even.s8   <- 0

mse.sd.even.mfvb <- 0
mse.sd.even.haar <- 0
mse.sd.even.s8   <- 0
mse.sd.even.j    <- 0
mse.sd.even.s8.j <- 0

lmse.sd.even.mfvb <- 0
lmse.sd.even.haar <- 0
lmse.sd.even.s8   <- 0
lmse.sd.even.j    <- 0
lmse.sd.even.s8.j <- 0

# Repeat for each data set simulated in the second simulation scenario.
cat(sprintf("Running %d simulations for Simulation Scenario 2.\n",nsim2))
for (j in 1:nsim2) {
  cat(sprintf("%d ",j))

  # Create the simulated data set
  # -----------------------------
  n      <- 2^10
  xOrig  <- (1:n)/n
  set.seed(30*j)
  yOrig  <- fTrue(xOrig) + sqrt(exp(loggTrue(xOrig)))*rnorm(n)
  aOrig  <- min(xOrig)
  bOrig  <- max(xOrig)
  mean.x <- mean(xOrig)
  sd.x   <- sd(xOrig)
  mean.y <- mean(yOrig)
  sd.y   <- sd(yOrig)

  a <- (aOrig - mean.x)/sd.x
  b <- (bOrig - mean.x)/sd.x
  x <- (xOrig - mean.x)/sd.x
  y <- (yOrig - mean.y)/sd.y

  numIntKnotsU <- 17
  intKnotsU <- quantile(x,seq(0,1,length=numIntKnotsU+2)[-c(1,numIntKnotsU+2)])
  Zu        <- ZOSull(x,intKnots=intKnotsU,range.x=c(a,b))
  numKnotsU <- ncol(Zu)

  numIntKnotsV <- numIntKnotsU 
  intKnotsV <- quantile(x,seq(0,1,length=numIntKnotsV+2)[-c(1,numIntKnotsV+2)])
  Zv        <- ZOSull(x,intKnots=intKnotsV,range.x=c(a,b))
  numKnotsV <- ncol(Zv)

  # Run Mean Field Variational Bayes (MFVB)
  # ---------------------------------------
  X     <- cbind(rep(1,n),x)
  Cumat <- cbind(X,Zu)
  Cvmat <- cbind(X,Zv)
  ncX   <- ncol(X)
  ncZu  <- ncol(Zu)
  ncZv  <- ncol(Zv)
  ncCu  <- ncol(Cumat)
  ncCv  <- ncol(Cvmat)

  MFVBfit <- meanVarMFVB(y,X,ncZu,ncZv,Au.hyp,Av.hyp,
                       sigsq.gamma,sigsq.beta)

  ng     <- 2^10
  xgOrig <- seq(aOrig,bOrig,length=ng)
  xg  <- (xgOrig - mean.x)/sd.x
  Xg  <- cbind(rep(1,ng),xg)
  Zug <- ZOSull(xg,intKnots=intKnotsU,range.x=c(a,b))
  Cug <- cbind(Xg,Zug)
  Zvg <- ZOSull(xg,intKnots=intKnotsV,range.x=c(a,b))
  Cvg <- cbind(Xg,Zvg)

  mu.q.nu       <- MFVBfit$mu.q.nu
  mu.q.omega    <- MFVBfit$mu.q.omega
  Sigma.q.nu    <- MFVBfit$Sigma.q.nu
  Sigma.q.omega <- MFVBfit$Sigma.q.omega

  # Get the mean function estimate.
  fhatMFVBg     <- Cug %*% mu.q.nu
  fhatMFVBgOrig <- fhatMFVBg*sd.y + mean.y

  logghatMFVBg     <- Cvg%*%mu.q.omega 
  logghatMFVBgOrig <- logghatMFVBg + 2*log(sd.y)

  sdloggMFVBgOrig      <- sqrt(diag(Cvg%*%Sigma.q.omega%*%t(Cvg))) 
  credLowloggMFVBgOrig <- logghatMFVBgOrig - qnorm(0.975)*sdloggMFVBgOrig
  credUpploggMFVBgOrig <- logghatMFVBgOrig + qnorm(0.975)*sdloggMFVBgOrig

  sqrtghatMFVBg     <- exp(0.5*Cvg%*%mu.q.omega 
                           + 0.125*diag(Cvg%*%Sigma.q.omega%*%t(Cvg)))
  sqrtghatMFVBgOrig <- sqrtghatMFVBg*sd.y

  # Run SMASH
  # ---------
  mu.est  <- smash.gaus(yOrig,filter.number=1,family="DaubExPhase")
  var.est <- smash.gaus(yOrig,v.est=TRUE)

  mse.mu.even.mfvb[j] <- mean((fhatMFVBgOrig-fTrue(xgOrig))^2)
  mse.mu.even.haar[j] <- mean((mu.est-fTrue(xgOrig))^2)

  mse.sd.even.mfvb[j]  <- mean((sqrtghatMFVBgOrig-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.even.haar[j]  <- mean((sqrt(var.est)-exp((loggTrue(xgOrig))/2))^2)
  lmse.sd.even.mfvb[j] <- mean((log(sqrtghatMFVBgOrig)-(loggTrue(xgOrig))/2)^2)
  lmse.sd.even.haar[j] <- mean((log(sqrt(var.est))-(loggTrue(xgOrig))/2)^2)

  mu.est       <- smash.gaus(yOrig,filter.number=8,family="DaubLeAsymm")
  var.est      <- smash.gaus(yOrig,v.est=TRUE,v.basis=TRUE,filter.number=8,
                             family="DaubLeAsymm")
  var.est.s8.j <- smash.gaus(yOrig,v.est=TRUE,v.basis=TRUE,jash=TRUE,
                             filter.number=8,family="DaubLeAsymm")
  var.est.j    <- smash.gaus(yOrig,v.est=TRUE,jash=TRUE)

  mse.mu.even.s8[j]   <-mean((mu.est-fTrue(xgOrig))^2)
  mse.sd.even.s8[j]   <-mean((sqrt(var.est)-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.even.j[j]    <-mean((sqrt(var.est.j)-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.even.s8.j[j] <-mean((sqrt(var.est.s8.j)-exp((loggTrue(xgOrig))/2))^2)
  lmse.sd.even.s8[j]  <-mean((log(sqrt(var.est))-(loggTrue(xgOrig))/2)^2)
  lmse.sd.even.j[j]   <-mean((log(sqrt(var.est.j))-(loggTrue(xgOrig))/2)^2)
  lmse.sd.even.s8.j[j]<-mean((log(sqrt(var.est.s8.j))-(loggTrue(xgOrig))/2)^2)
}
cat("\n")

# PLOT MEAN & VARIANCE FUNCTIONS USED TO SIMULATE DATA
# ----------------------------------------------------
# Compare this plot against the one shown in Fig. 4 of the paper.
cat("Plotting mean and variance functions used to simulate data.\n")
xgrid <- (0:10000)/10000
plot(xgrid,fTrue(xgrid),type = "l",ylim = c(-5,5),ylab = "y",xlab = "X",
     lwd = 2)
lines(xgrid,fTrue(xgrid) + 2*sqrt(gTrue(xgrid)),col = "darkorange",lwd = 2)
lines(xgrid,fTrue(xgrid) - 2*sqrt(gTrue(xgrid)),col = "darkorange",lwd = 2)

# SUMMARIZE RESULTS
# -----------------
# Compare this table against Table 1 in the paper.
cat("Summarizing results of simulations.\n")
cat(sprintf("MSE averaged across %d simulations in Scenario 1,\n",nsim1))
cat(sprintf("and averaged across %d simulations in Scenario 2:\n",nsim2))
mse.table <- rbind(c(mean(mse.mu.uneven.mfvb),mean(mse.sd.uneven.mfvb),
                     mean(mse.mu.even.mfvb),mean(mse.sd.even.mfvb)),
                   c(mean(mse.mu.uneven.s8),mean(mse.sd.uneven.s8),
                     mean(mse.mu.even.s8),mean(mse.sd.even.s8)))
rownames(mse.table) <- c("MFVB","SMASH")
colnames(mse.table) <- c("mean","sd","mean","sd")
cat("         Scenario 1      Scenario 2   \n")
cat("      --------------- ----------------\n")
print(mse.table)
```