### 2018-12-07 analysis/gauss_shiny_setup.Rmd
Revised the Shiny plot; now I need to provide instructions for recreating it.

---
title: "Instructions for running Shiny server to explore results of Gaussian mean estimation experiments"
author: "Zhengrong Xing, Peter Carbonetto and Matthew Stephens"
output: workflowr::wflow_html
---

*Briefly explain here what this R Markdown document is for.*

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE)
```

```{r load-pkgs, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(shiny)
library(rmarkdown)
```

```{r load-results, eval=FALSE}
load("../output/dscr.RData")
```

NOTE: To launch the Shiny app, run:

```{r launch-shiny eval=FALSE}
run("../shiny/gauss_shiny.Rmd")
```

The following code generates an RShiny app which displays boxplots of
MISEs from the full simulation study comparing SMASH to other
wavelet-based methods for denoising Gaussian data. As mentioned in the
companion paper, the results for a variety of test functions (both
mean and variance functions) and 2 signal-to-noise ratios (SNRs), 1
and 3, are presented here.