### 2018-12-18 analysis/gauss_shiny_setup.Rmd
Added link to Shiny package.

---
title: "Instructions for running Shiny app to explore results of Gaussian mean estimation experiments"
author: "Zhengrong Xing, Peter Carbonetto and Matthew Stephens"
output: workflowr::wflow_html
---

Here we give instructions for starting the [Shiny app][shiny] which
will create an interactive plot from the results of the Gaussian mean
estimation simulation study.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE)
```

First, you need make sure that the following R packages are installed
on your computer: dplyr, ggplot2, cowplot, shiny and rmarkdown. See
the "Session Info" at the bottom for the versions of the R packages we
used.  Once you have installed these packages, run this code to verify
that the packages are installed successfully:

```{r load-pkgs, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(cowplot)
library(shiny)
library(rmarkdown)
```

Now you are ready to launch the Shiny app with this command:

```{r launch-shiny, eval=FALSE}
rmarkdown::run("../shiny/gauss_shiny.Rmd")
```

The interactive plot will be launched in your Web browser, or in an
RStudio panel, depending on whether you are using R or RStudio.

[shiny]: https://cran.r-project.org/package=shiny