### 2018-12-19 analysis/poisson_tables.Rmd

---
title: "Title goes here."
author: "Zhengrong Xing, Peter Carbonetto and Matthew Stephens"
output: workflowr::wflow_html
---

This script produces supplementary tables for Poisson simulations.

```{r}
library(xtable)

load("../output/pois.RData")

tex.row.names <- c("SMASH",
                   "BMSM",
                   "BMMIM",
                   "Haar-Fisz",
                   "Anscombe",
                   "Haar thresholds",
                   "L1_penalty"
)

tex.col.names <- c("intensity (0.01,3)",
                   "intensity (1/8,8)",
                   "intensity (1/128,128)")

mise.s.1.table=mise.s.1[c(1,2,3,4,6,7,8)]
mise.s.8.table=mise.s.8[c(1,2,3,4,6,7,8)]
mise.s.128.table=mise.s.128[c(1,2,3,4,6,7,8)]

mise.ang.1.table=mise.ang.1[c(1,2,3,4,6,7,8)]
mise.ang.8.table=mise.ang.8[c(1,2,3,4,6,7,8)]
mise.ang.128.table=mise.ang.128[c(1,2,3,4,6,7,8)]

mise.bur.1.table=mise.bur.1[c(1,2,3,4,6,7,8)]
mise.bur.8.table=mise.bur.8[c(1,2,3,4,6,7,8)]
mise.bur.128.table=mise.bur.128[c(1,2,3,4,6,7,8)]

mise.cb.1.table=mise.cb.1[c(1,2,3,4,6,7,8)]
mise.cb.8.table=mise.cb.8[c(1,2,3,4,6,7,8)]
mise.cb.128.table=mise.cb.128[c(1,2,3,4,6,7,8)]

mise.b.1.table=mise.b.1[c(1,2,3,4,6,7,8)]
mise.b.8.table=mise.b.8[c(1,2,3,4,6,7,8)]
mise.b.128.table=mise.b.128[c(1,2,3,4,6,7,8)]

mise.hs.1.table=mise.hs.1[c(1,2,3,4,6,7,8)]
mise.hs.8.table=mise.hs.8[c(1,2,3,4,6,7,8)]
mise.hs.128.table=mise.hs.128[c(1,2,3,4,6,7,8)]

mise.s.table=cbind(mise.s.1.table,
mise.s.8.table,
mise.s.128.table
)

mise.ang.table <- cbind(mise.ang.1.table,
mise.ang.8.table,
mise.ang.128.table
)

mise.bur.table=cbind(mise.bur.1.table,
mise.bur.8.table,
mise.bur.128.table
)

mise.cb.table=cbind(mise.cb.1.table,
mise.cb.8.table,
mise.cb.128.table
)

mise.b.table=cbind(mise.b.1.table,
mise.b.8.table,
mise.b.128.table
)

mise.hs.table=cbind(mise.hs.1.table,
mise.hs.8.table,
mise.hs.128.table
)

rownames(mise.ang.table)=tex.row.names
colnames(mise.ang.table)=tex.col.names
rownames(mise.b.table)=tex.row.names
colnames(mise.b.table)=tex.col.names
rownames(mise.cb.table)=tex.row.names
colnames(mise.cb.table)=tex.col.names
rownames(mise.s.table)=tex.row.names
colnames(mise.s.table)=tex.col.names
rownames(mise.hs.table)=tex.row.names
colnames(mise.hs.table)=tex.col.names
rownames(mise.bur.table)=tex.row.names
colnames(mise.bur.table)=tex.col.names
```

```{r, results="asis"}
print(xtable(mise.s.table,caption="Comparison of methods for denoising Poisson data for the ``Spikes'' test function for 3 different (min,max) intensities ((0.01,3), (1/8,8), (1/128,128)). Performance is measured using MISE over 100 independent datasets, with smaller values indicating better performance. Values colored in red indicates the smallest MISE amongst all methods for a given (min, max) intensity.",label="table:pois_sp",digits=2),type = "html")
```

```{r, results="asis"}
print(xtable(mise.ang.table,caption="Comparison of methods for denoising Poisson data for the ``Angles'' test function for 3 different (min,max) intensities ((0.01,3), (1/8,8), (1/128,128)). Performance is measured using MISE over 100 independent datasets, with smaller values indicating better performance. Values colored in red indicates the smallest MISE amongst all methods for a given (min, max) intensity.",label="table:pois_ang",digits=2),type = "html")
```

```{r, results="asis"}
print(xtable(mise.hs.table,caption="Comparison of methods for denoising Poisson data for the ``Heavisine'' test function for 3 different (min,max) intensities ((0.01,3), (1/8,8), (1/128,128)). Performance is measured using MISE over 100 independent datasets, with smaller values indicating better performance. Values colored in red indicates the smallest MISE amongst all methods for a given (min, max) intensity.",label="table:pois_hs",digits=2),type = "html")
```

```{r, results="asis"}
print(xtable(mise.bur.table,caption="Comparison of methods for denoising Poisson data for the ``Bursts'' test function for 3 different (min,max) intensities ((0.01,3), (1/8,8), (1/128,128)). Performance is measured using MISE over 100 independent datasets, with smaller values indicating better performance. Values colored in red indicates the smallest MISE amongst all methods for a given (min, max) intensity.",label="table:pois_bur",digits=2),type = "html")
```

```{r, results="asis"}
print(xtable(mise.cb.table,caption="Comparison of methods for denoising Poisson data for the ``Clipped Blocks'' test function for 3 different (min,max) intensities ((0.01,3), (1/8,8), (1/128,128)). Performance is measured using MISE over 100 independent datasets, with smaller values indicating better performance. Values colored in red indicates the smallest MISE amongst all methods for a given (min, max) intensity.",label="table:pois_cb",digits=2),type = "html")
```

```{r, results="asis"}
print(xtable(mise.b.table,caption="Comparison of methods for denoising Poisson data for the ``Bumps'' test function for 3 different (min,max) intensities ((0.01,3), (1/8,8), (1/128,128)). Performance is measured using MISE over 100 independent datasets, with smaller values indicating better performance. Values colored in red indicates the smallest MISE amongst all methods for a given (min, max) intensity.",label="table:pois_b",digits=2),type = "html")
```