### 2019-11-08 analysis/gausdemo.Rmd
Added data points to gausdemo plots.

This demo plots SMASH against TI-thresholding (using RMAD to estimate variance) for a variety of mean-variance-snr combinations.

```{r demo, fig.height=4, fig.width=6}
library(AlgDesign)
library(smashr)
source("../code/signals.R")

# today's date
set.seed(191027)

meanfn = c(spikes.fn, bumps.fn, blocks.fn, angles.fn, doppler.fn, blip.fn, cor.fn)
meanfn.short = c("sp", "bump", "blk", "ang", "dop", "blip", "cor")
rsnr = sqrt(c(1, 3))
varfn = c(cons.fn, texp.fn, doppler.fn, cblocks.fn)
varfn.short = c("cons", "texp", "dop", "cblk")

design = gen.factorial(c(length(varfn), length(rsnr), length(meanfn)), center = FALSE)

for (i in 1:dim(design)[1]) {
  name = paste(meanfn.short[design[i, 3]], rsnr[design[i, 2]]^2, varfn.short[design[i, 1]], sep = ".")
  args = list(n = 1024, rsnr = rsnr[design[i, 2]], meanfn = meanfn[[design[i, 3]]], varfn = varfn[[design[i, 1]]])

  t = (1:1024)/1024

  mu = args$meanfn(t, "mean")
  sigma = sqrt(args$varfn(t, "var"))
  sd = sigma/mean(sigma) * sd(mu)/rsnr^2
  x.sim = rnorm(args$n, mu, sd)

  mu.smash = smash(x.sim,family = "DaubLeAsymm",filter.number = 8)
  mu.ti = ti.thresh(x.sim,method = "rmad",family = "DaubLeAsymm",
                        filter.number = 8)
  title = paste("mean: ", meanfn.short[design[i, 3]], ", var: ", varfn.short[design[i, 1]], ", snr: ", rsnr[design[i, 2]]^2, sep = "")
  par(cex.axis = 1)
  plot(mu,type = "l",col = "black",lwd = 3,xlab = "position",ylab = "",
       ylim = c(-0.05,1),xaxp = c(0,1024,4),yaxp = c(0,1,4), main = title)
  points(x.sim,cex = 0.7,pch = 1,col = "skyblue")
  lines(mu.ti,col = "dodgerblue",lwd = 3)
  lines(mu.smash,col = "orangered",lwd = 3)
  legend(650, 1.05, c("ground-truth signal", "TI-thresholding (RMAD)", "SMASH"), col = c("black", "dodgerblue", "orangered"), bty = "n", lwd = 1, pt.cex = 1, cex = 0.75)
}
```