### 2019-11-12 analysis/gaussian_signals.Rmd
wflow_publish("gaussian_signals.Rmd")

---
title: "Mean and variance functions used to simulate Gaussian data"
author: "Zhengrong Xing, Peter Carbonetto and Matthew Stephens"
output: workflowr::wflow_html
---

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,fig.align = "center")
```

Set up environment
------------------

Load the ggplot2 and cowplot packages, and the functions definining
the mean and variances used to simulate the data.

```{r load-pkgs, message=FALSE, warning=FALSE}
library(ggplot2)
library(cowplot)
source("../code/signals.R")
```

Generate the ground-truth signals
---------------------------------

Here, `n` specifies the length of the signals.

```{r specify-n}
n <- 1024
t <- 1:n/n
```

Define the Angle mean function.

```{r define-angle} 
sig = ((2 * t + 0.5) * (t <= 0.15)) +
      ((-12 * (t - 0.15) + 0.8) * (t > 0.15 & t <= 0.2)) +
      0.2 * (t > 0.2 & t <= 0.5) + 
      ((6 * (t - 0.5) + 0.2) * (t > 0.5 & t <= 0.6)) +
      ((-10 * (t - 0.6) + 0.8) * (t > 0.6 & t <= 0.65)) +
      ((-0.5 * (t - 0.65) + 0.3) * (t > 0.65 & t <= 0.85)) +
      ((2 * (t - 0.85) + 0.2) * (t > 0.85))
mu.ang = 3/5 * ((5/(max(sig) - min(sig))) * sig - 1.6) - 0.0419569
```
Define the Block mean and variance functions.

```{r define-block}
pos    = c(0.1, 0.13, 0.15, 0.23, 0.25, 0.4, 0.44, 0.65, 0.76, 0.78, 0.81)
hgt    = 2.88/5 * c(4, (-5), 3, (-4), 5, (-4.2), 2.1, 4.3, (-3.1), 2.1, (-4.2))
mu.blk = rep(0, n)
for (j in 1:length(pos))
  mu.blk = mu.blk + (1 + sign(t - pos[j])) * (hgt[j]/2)
mu.cblk = mu.blk
mu.cblk[mu.cblk < 0] = 0
```

Define the mean functions.

```{r define-mean}
mu.sp   <- spike.fn(t,"mean")
mu.bump <- bumps.fn(t,"mean")
mu.dop  <- doppler.fn(t,"mean")
mu.blip <- blip.fn(t,"mean")
mu.cor  <- cor.fn(t,"mean")
```

Define the remainder of the mean functions.

```{r define-mean-2}
mu.blk  = 0.2 + 0.6 * (mu.blk - min(mu.blk))/max(mu.blk - min(mu.blk))
mu.ang  = (1 + mu.ang)/5
```

Define the variance functions.

```{r define-variance}
var1 = rep(1, n)
var2 = (0.01 + 4 * (exp(-550 * (t - 0.2)^2) +
                     exp(-200 * (t - 0.5)^2) +
                     exp(-950 * (t - 0.8)^2)))
var3 <- doppler.fn(t,"var")
var4 <- bumps.fn(t,"var")
var5 = 0.01 + 1 * (mu.cblk - min(mu.cblk))/max(mu.cblk)
var1 = var1/2
var2 = var2/max(var2)
var5 = var5/max(var5)
```

Plot the signal means
---------------------

This function will be used to draw the mean and variance functions.

```{r define-plot-signal}
plot.signal <- function (t, y, label)
  quickplot(t,y,geom = "line",color = I("darkorange"),
            xlab = "",ylab = "",main = label)
```

These plots show each of the mean functions used in generating the
Gaussian data sets.

```{r plot-mean-functions, fig.height=10, fig.width=6}
theme_set(theme_cowplot())
plot_grid(plot.signal(t,mu.sp,"Spikes (sp)"),
          plot.signal(t,mu.bump,"Bumps (bump)"),
          plot.signal(t,mu.blk,"Blocks (blk)"),
          plot.signal(t,mu.ang,"Angles (ang)"),
          plot.signal(t,mu.dop,"Doppler (dop)"),
          plot.signal(t,mu.blip,"Blip (blip)"),
          plot.signal(t,mu.cor,"Corner (cor)"),
		  nrow = 4,ncol = 2)
```

Plot the signal variances
-------------------------

These plots show the variance functions used in generating the
Gaussian data sets. In practice, these functions are rescaled in the
simulations to achieve the desired signal-to-noise ratios (see the
paper for a more detailed explanation).

```{r plot-variance-functions, fig.height=7.5, fig.width=6}
plot_grid(plot.signal(t,var1,"Constant variance (v1)"),
          plot.signal(t,var2,"Triple exponential (v2)"),
          plot.signal(t,var3,"Doppler (v3)"),
		  plot.signal(t,var4,"Bumps (v4)"),
		  plot.signal(t,var5,"Clipped (v5)"),
		  nrow = 3,ncol = 2)
```