### 2019-11-12 analysis/gaussian_signals.Rmd
wflow_publish("gaussian_signals.Rmd")

---
title: "Mean and variance functions used to simulate Gaussian data"
author: "Zhengrong Xing, Peter Carbonetto and Matthew Stephens"
output: workflowr::wflow_html
---

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,fig.align = "center")
```

Set up environment
------------------

Load the ggplot2 and cowplot packages, and the functions definining
the mean and variances used to simulate the data.

```{r load-pkgs, message=FALSE, warning=FALSE}
library(ggplot2)
library(cowplot)
source("../code/signals.R")
```

Generate the ground-truth signals
---------------------------------

Here, `n` specifies the length of the signals.

```{r specify-n}
n <- 1024
t <- 1:n/n
```

Define the mean functions.

```{r define-mean}
mu.sp   <- spike.fn(t,"mean")
mu.bump <- bumps.fn(t,"mean")
mu.blk  <- blocks.fn(t,"mean")
mu.ang  <- angles.fn(t,"mean")
mu.dop  <- doppler.fn(t,"mean")
mu.blip <- blip.fn(t,"mean")
mu.cor  <- cor.fn(t,"mean")
```

Define the variance functions.

```{r define-variance}
var1 <- cons.fn(t,"var")
var2 <- texp.fn(t,"var")
var3 <- doppler.fn(t,"var")
var4 <- bumps.fn(t,"var")
var5 <- cblocks.fn(t,"var")
```

Plot the signal means
---------------------

This function is used to draw the mean and variance functions.

```{r define-plot-signal}
plot.signal <- function (t, y, label)
  quickplot(t,y,geom = "line",color = I("darkorange"),
            xlab = "",ylab = "",main = label)
```

These plots show each of the mean functions used in generating the
Gaussian data sets.

```{r plot-mean-functions, fig.height=10, fig.width=6}
theme_set(theme_cowplot())
plot_grid(plot.signal(t,mu.sp,"Spikes (sp)"),
          plot.signal(t,mu.bump,"Bumps (bump)"),
          plot.signal(t,mu.blk,"Blocks (blk)"),
          plot.signal(t,mu.ang,"Angles (ang)"),
          plot.signal(t,mu.dop,"Doppler (dop)"),
          plot.signal(t,mu.blip,"Blip (blip)"),
          plot.signal(t,mu.cor,"Corner (cor)"),
		  nrow = 4,ncol = 2)
```

Plot the signal variances
-------------------------

These plots show the variance functions used in generating the
Gaussian data sets. In practice, these functions are rescaled in the
simulations to achieve the desired signal-to-noise ratios (see the
paper for a more detailed explanation).

```{r plot-variance-functions, fig.height=7.5, fig.width=6}
plot_grid(plot.signal(t,var1,"Constant variance (v1)"),
          plot.signal(t,var2,"Triple exponential (v2)"),
          plot.signal(t,var3,"Doppler (v3)"),
		  plot.signal(t,var4,"Bumps (v4)"),
		  plot.signal(t,var5,"Clipped (v5)"),
		  nrow = 3,ncol = 2)
```