### 2020-06-16 analysis/gaussmeanest.Rmd
wflow_publish("gaussmeanest.Rmd")

---
title: "Gaussian mean estimation in simulated data sets"
author: "Zhengrong Xing, Peter Carbonetto and Matthew Stephens"
output: workflowr::wflow_html
---

In this analysis, we assess the ability of different signal denoising
methods to recover the true signal after being provided with
Gaussian-distributed observations of the signal. We consider scenarios
in which the data have homoskedastic errors (constant variance) and
heteroskedastic errors (non-constant variance).

Since the simulation experiments are computationally intensive, here
we only illustrate the application of the signal denoising methods,
and create plots summarizing the results of the full experiments; the
full experiments were implemented separately. (For instructions on
re-running these simulation experiments, see the README in the "dsc"
directory of the git repository).

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,fig.align = "center")
```

Set up environment
------------------

Load the ggplot2 and cowplot packages, and the functions definining
the mean and variances used to simulate the data.

```{r load-pkgs, message=FALSE, warning=FALSE}
library(plyr)
library(smashr)
library(ggplot2)
library(cowplot)
source("../code/signals.R")
source("../code/gaussmeanest.functions.R")
theme_set(theme_cowplot())
```

Load results
------------

Load the results of the simulation experiments.

```{r load-results, warning=FALSE}
load("../output/dscr.RData")
```

Simulated data with constant variances
--------------------------------------

This plot reproduces Fig. 2 of the manuscript, which compares the
accuracy of the mean curves estimated from the data sets that were
simulated using the "Spikes" mean function with constant variance and
a signal-to-noise ratio of 3.

First, extract the results used to generate this plot, and transform
them into a data frame suitable for plotting using ggplot2.

```{r homo-extract-data}
pdat <- get.results.homosked(res,"sp.3.v1")
```

Create the combined boxplot and violin plot using ggplot2.

```{r create-violin-plot-homo, fig.height=4, fig.width=5}
pdat <-
  transform(pdat,
            method = factor(method,
                            names(sort(tapply(pdat$mise,pdat$method,mean),
                                       decreasing = TRUE))))
p <- ggplot(pdat,aes(x = method,y = mise,fill = method.type)) +
     geom_violin(fill = "skyblue",color = "skyblue") +
     geom_boxplot(width = 0.15,outlier.shape = NA) +
     scale_y_continuous(breaks = seq(6,16,2)) +
     scale_fill_manual(values = c("darkorange","dodgerblue","gold"),
                       guide = FALSE) +
     coord_flip() +
     labs(x = "",y = "MISE") +
     theme(axis.line = element_blank(),
           axis.ticks.y = element_blank())
print(p)
```

From this plot, we see that the three variants of SMASH all
outperformed EbayesThresh and TI thresholding in this setting.

These plots summarize the results for all 7 simulation scenarios and
the two signal-to-noise ratios (1 and 3), including the "Spikes"
scenario shown in greater detail in the violin plot above.

```{r summarize-results-homo, fig.height=4, fig.width=8}
create.bar.plots.homosked(rbind(
  data.frame(sim = "sp",  snr = 1,get.results.homosked(res,"sp.1.v1")),
  data.frame(sim = "bump",snr = 1,get.results.homosked(res,"bump.1.v1")),
  data.frame(sim = "blkp",snr = 1,get.results.homosked(res,"blk.1.v1")),
  data.frame(sim = "ang", snr = 1,get.results.homosked(res,"ang.1.v1")),
  data.frame(sim = "dop", snr = 1,get.results.homosked(res,"dop.1.v1")),
  data.frame(sim = "blip",snr = 1,get.results.homosked(res,"blip.1.v1")),
  data.frame(sim = "cor", snr = 1,get.results.homosked(res,"cor.1.v1")),
  data.frame(sim = "sp",  snr = 3,get.results.homosked(res,"sp.3.v1")),
  data.frame(sim = "bump",snr = 3,get.results.homosked(res,"bump.3.v1")),
  data.frame(sim = "blkp",snr = 3,get.results.homosked(res,"blk.3.v1")),
  data.frame(sim = "ang", snr = 3,get.results.homosked(res,"ang.3.v1")),
  data.frame(sim = "dop", snr = 3,get.results.homosked(res,"dop.3.v1")),
  data.frame(sim = "blip",snr = 3,get.results.homosked(res,"blip.3.v1")),
  data.frame(sim = "cor", snr = 3,get.results.homosked(res,"cor.3.v1"))))
```

Next, we compare the same methods in simulated data sets with
heteroskedastic errors.

Simulated data with heteroskedastic errors: "Spikes" mean signal and "Clipped Blocks" variance
----------------------------------------------------------------------------------------------

In this scenario, the data sets were simulated using the "Spikes" mean
function and the "Clipped Blocks" variance function. The next two
plots reproduce part of Fig. 3 in the manuscript.

This plot shows the mean function as an orange line, the +/- 2
standard deviations as dashed, light blue lines, and an example
simulated data set (black circles):

```{r spikes-signal-with-data, fig.height=4, fig.width=6}
t         <- (1:1024)/1024
mu        <- spike.fn(t,"mean")
sigma.ini <- sqrt(cblocks.fn(t,"var"))
sd.fn     <- sigma.ini/mean(sigma.ini) * sd(mu)/3
x.sim     <- rnorm(1024,mu,sd.fn)
par(cex.axis = 1,cex.lab = 1.25)
plot(mu,type = "l", ylim = c(-0.05,1),col = "darkorange",
     xlab = "position",ylab = "",lwd = 1.75,xaxp = c(0,1024,4),
     yaxp = c(0,1,4))
lines(mu + 2*sd.fn,col = "dodgerblue",lty = "dashed",lwd = 1.75)
lines(mu - 2*sd.fn,col = "dodgerblue",lty = "dashed",lwd = 1.75)
points(x.sim,cex = 0.7,pch = 1,col = "black")
```

Next, we plot the ground-truth signal (the mean function, drawn as a
black line), and the signals recovered by TI thresholding (light blue
line) and SMASH (the red line) for the example simulated data set.

```{r spikes-signal-with-estimates, fig.height=4, fig.width=6}
mu.smash <- smash(x.sim,family = "DaubLeAsymm",filter.number = 8)
mu.ti    <- ti.thresh(x.sim,method = "rmad",family = "DaubLeAsymm",
                      filter.number = 8)
par(cex.axis = 1)
plot(mu,type = "l",col = "black",lwd = 3,xlab = "position",ylab = "",
     ylim = c(-0.05,1),xaxp = c(0,1024,4),yaxp = c(0,1,4))
lines(mu.ti,col = "dodgerblue",lwd = 3)
lines(mu.smash,col = "orangered",lwd = 3)
```

Extract the results from running the simulations.

```{r plot-2-extract-data}
hetero.data.smash <-
  res[res$.id == "sp.3.v5" & res$method == "smash.s8",]
hetero.data.smash.homo <-
  res[res$.id == "sp.3.v5" & res$method == "smash.homo.s8",]
hetero.data.tithresh.homo <-
  res[res$.id == "sp.3.v5" & res$method == "tithresh.homo.s8",]
hetero.data.tithresh.rmad <-
  res[res$.id == "sp.3.v5" & res$method == "tithresh.rmad.s8",]
hetero.data.tithresh.smash <-
  res[res$.id == "sp.3.v5" & res$method == "tithresh.smash.s8",]
hetero.data.tithresh.true <-
  res[res$.id == "sp.3.v5" & res$method == "tithresh.true.s8",]
hetero.data.ebayes <-
  res[res$.id == "sp.3.v5" & res$method == "ebayesthresh",]
hetero.data.smash.true <-
  res[res$.id == "sp.3.v5" & res$method == "smash.true.s8",]
```

Transform these results into a data frame suitable for ggplot2.

```{r plot-2-transform-data}
pdat <-
  rbind(data.frame(method      = "smash",
                   method.type = "est",
                   mise        = hetero.data.smash$mise),
        data.frame(method      = "smash.homo",
                   method.type = "homo",
                   mise        = hetero.data.smash.homo$mise),
        data.frame(method      = "tithresh.rmad",
                   method.type = "tithresh",
                   mise        = hetero.data.tithresh.rmad$mise),
        data.frame(method      = "tithresh.smash",
                   method.type = "tithresh",
                   mise        = hetero.data.tithresh.smash$mise),
        data.frame(method      = "tithresh.true",
                   method.type = "tithresh",
                   mise        = hetero.data.tithresh.true$mise),
        data.frame(method      = "ebayesthresh",
                   method.type = "homo",
                   mise        = hetero.data.ebayes$mise),
        data.frame(method      = "smash.true",
                   method.type = "true",
                   mise        = hetero.data.smash.true$mise))
pdat <-
  transform(pdat,
            method = factor(method,
                            names(sort(tapply(pdat$mise,pdat$method,mean),
                                       decreasing = TRUE))))
```

Create the combined boxplot and violin plot using ggplot2.

```{r plot-2-create, fig.height=4, fig.width=5}
p <- ggplot(pdat,aes(x = method,y = mise,fill = method.type)) +
     geom_violin(fill = "skyblue",color = "skyblue") +
     geom_boxplot(width = 0.15,outlier.shape = NA) +
     scale_fill_manual(values=c("darkorange","dodgerblue","limegreen","gold"),
                       guide = FALSE) +
     coord_flip() +
     scale_y_continuous(breaks = seq(10,70,10)) +
     labs(x = "",y = "MISE") +
     theme(axis.line = element_blank(),
           axis.ticks.y = element_blank())
print(p)
```

In this scenario, we see that SMASH, when allowing for heteroskedastic
errors, outperforms EbayesThresh and all variants of TI thresholding
(including TI thresholding when provided with the true variance).
Further, SMASH performs almost as well when estimating the variance
compared to when provided with the true variance.

Simulated data with heteroskedastic errors: "Corner" mean signal and "Doppler" variance
---------------------------------------------------------------------------------------

In this next scenario, the data sets were simulated using the "Corner"
mean function and the "Doppler" variance function. These plots were
also used in Figure 3 of the manuscript.

This plot shows the mean function as an orange line, the +/- 2
standard deviations as dashed, light blue lines, and an example
simulated data set (black circles):

```{r corner-signal, fig.height=4, fig.width=6}
t         <- (1:1024)/1024
mu        <- cor.fn(t,"mean") 
sigma.ini <- sqrt(doppler.fn(t,"var"))
sd.fn     <- sigma.ini/mean(sigma.ini) * sd(mu)/3
x.sim     <- rnorm(1024,mu,sd.fn)
plot(mu,type = "l",col = "darkorange",ylim = c(-0.05,1),
     xlab = "position",ylab = "",lwd = 1.75,xaxp = c(0,1024,4),
     yaxp = c(0,1,4))
lines(mu + 2*sd.fn,col = "dodgerblue",lty = "dashed",lwd = 1.75)
lines(mu - 2*sd.fn,col = "dodgerblue",lty = "dashed",lwd = 1.75)
points(x.sim,cex = 0.7,pch = 1,col = "black")
```

Now, we plot the ground-truth signal (the mean function, drawn as
a black line) and the signals recovered by TI thresholding (light blue
line) and SMASH (the red line) for one simulated dataset as an illustration

```{r corner-signal-with-estimates, fig.height=4, fig.width=6}
mu.smash <- smash(x.sim,family = "DaubLeAsymm",filter.number = 8)
mu.ti    <- ti.thresh(x.sim,method = "rmad",family = "DaubLeAsymm",
                      filter.number = 8)
par(cex.axis = 1)
plot(mu,type = "l",col = "black",lwd = 3,xlab = "position",ylab = "",
     ylim = c(-0.05,1),xaxp = c(0,1024,4),yaxp = c(0,1,4))
lines(mu.ti,col = "dodgerblue",lwd = 3)
lines(mu.smash,col = "orangered",lwd = 3)
```

Extract the results from running these simulations.

```{r plot-3-extract-data}
hetero.data.smash.2 <-
  res[res$.id == "cor.3.v3" & res$method == "smash.s8",]
hetero.data.smash.homo.2 <-
  res[res$.id == "cor.3.v3" & res$method == "smash.homo.s8",]
hetero.data.tithresh.homo.2 <-
  res[res$.id == "cor.3.v3" & res$method == "tithresh.homo.s8",]
hetero.data.tithresh.rmad.2 <-
  res[res$.id == "cor.3.v3" & res$method == "tithresh.rmad.s8",]
hetero.data.tithresh.smash.2 <-
  res[res$.id == "cor.3.v3" & res$method == "tithresh.smash.s8",]
hetero.data.tithresh.true.2 <-
  res[res$.id == "cor.3.v3" & res$method == "tithresh.true.s8",]
hetero.data.ebayes.2 <-
  res[res$.id == "cor.3.v3" & res$method == "ebayesthresh",]
hetero.data.smash.true.2 <-
  res[res$.id == "cor.3.v3" & res$method == "smash.true.s8",]
```

Transform these results into a data frame suitable for ggplot2.

```{r plot-3-transform-data}
pdat <-
  rbind(data.frame(method      = "smash",
                   method.type = "est",
                   mise        = hetero.data.smash.2$mise),
        data.frame(method      = "smash.homo",
                   method.type = "homo",
                   mise        = hetero.data.smash.homo.2$mise),
        data.frame(method      = "tithresh.rmad",
                   method.type = "tithresh",
                   mise        = hetero.data.tithresh.rmad.2$mise),
        data.frame(method      = "tithresh.smash",
                   method.type = "tithresh",
                   mise        = hetero.data.tithresh.smash.2$mise),
        data.frame(method      = "tithresh.true",
                   method.type = "tithresh",
                   mise        = hetero.data.tithresh.true.2$mise),
        data.frame(method      = "ebayesthresh",
                   method.type = "homo",
                   mise        = hetero.data.ebayes.2$mise),
        data.frame(method      = "smash.true",
                   method.type = "true",
                   mise        = hetero.data.smash.true.2$mise))
pdat <-
  transform(pdat,
            method = factor(method,
                            names(sort(tapply(pdat$mise,pdat$method,mean),
                                       decreasing = TRUE))))
```

Create the combined boxplot and violin plot using ggplot2.

```{r plot-3-create, fig.height=4, fig.width=5}
p <- ggplot(pdat,aes(x = method,y = mise,fill = method.type)) +
     geom_violin(fill = "skyblue",color = "skyblue") +
     geom_boxplot(width = 0.15,outlier.shape = NA) +
     scale_fill_manual(values=c("darkorange","dodgerblue","limegreen","gold"),
                       guide = FALSE) +
     coord_flip() +
     scale_y_continuous(breaks = seq(1,5)) +
     labs(x = "",y = "MISE") +
     theme(axis.line = element_blank(),
           axis.ticks.y = element_blank())
print(p)
```

Similar to the "Spikes" scenario, we see that the SMASH method, when
allowing for heteroskedastic variances, outperforms both the TI
thresholding and EbayesThresh approaches.

Combined results from simulated data sets with heteroskedastic errors
---------------------------------------------------------------------

These plots summarize the results from all combinations of test
functions and a signal-to-noise ratio of 1.

```{r summarize-results-heterosked-snr-1, fig.height=4.5, fig.width=9}
create.bar.plots.heterosked(rbind(
  data.frame(mean="sp",  var="v2",get.results.heterosked(res,"sp.1.v2")),
  data.frame(mean="sp",  var="v3",get.results.heterosked(res,"sp.1.v3")),
  data.frame(mean="sp",  var="v4",get.results.heterosked(res,"sp.1.v4")),
  data.frame(mean="sp",  var="v5",get.results.heterosked(res,"sp.1.v5")),
  data.frame(mean="bump",var="v2",get.results.heterosked(res,"bump.1.v2")),
  data.frame(mean="bump",var="v3",get.results.heterosked(res,"bump.1.v3")),
  data.frame(mean="bump",var="v4",get.results.heterosked(res,"bump.1.v4")),
  data.frame(mean="bump",var="v5",get.results.heterosked(res,"bump.1.v5")),
  data.frame(mean="blk", var="v2",get.results.heterosked(res,"blk.1.v2")),
  data.frame(mean="blk", var="v3",get.results.heterosked(res,"blk.1.v3")),
  data.frame(mean="blk", var="v4",get.results.heterosked(res,"blk.1.v4")),
  data.frame(mean="blk", var="v5",get.results.heterosked(res,"blk.1.v5")),
  data.frame(mean="ang", var="v2",get.results.heterosked(res,"ang.1.v2")),
  data.frame(mean="ang", var="v3",get.results.heterosked(res,"ang.1.v3")),
  data.frame(mean="ang", var="v4",get.results.heterosked(res,"ang.1.v4")),
  data.frame(mean="ang", var="v5",get.results.heterosked(res,"ang.1.v5")),
  data.frame(mean="dop", var="v2",get.results.heterosked(res,"dop.1.v2")),
  data.frame(mean="dop", var="v3",get.results.heterosked(res,"dop.1.v3")),
  data.frame(mean="dop", var="v4",get.results.heterosked(res,"dop.1.v4")),
  data.frame(mean="dop", var="v5",get.results.heterosked(res,"dop.1.v5")),
  data.frame(mean="blip",var="v2",get.results.heterosked(res,"blip.1.v2")),
  data.frame(mean="blip",var="v3",get.results.heterosked(res,"blip.1.v3")),
  data.frame(mean="blip",var="v4",get.results.heterosked(res,"blip.1.v4")),
  data.frame(mean="blip",var="v5",get.results.heterosked(res,"blip.1.v5")),
  data.frame(mean="cor", var="v2",get.results.heterosked(res,"cor.1.v2")),
  data.frame(mean="cor", var="v3",get.results.heterosked(res,"cor.1.v3")),
  data.frame(mean="cor", var="v4",get.results.heterosked(res,"cor.1.v4")),
  data.frame(mean="cor", var="v5",get.results.heterosked(res,"cor.1.v5"))))
```

These plots summarize the results from data sets simulated using a
signal-to-noise ratio of 3.

```{r summarize-results-heterosked-snr-3, fig.height=4.5, fig.width=9}
create.bar.plots.heterosked(rbind(
  data.frame(mean="sp",  var="v2",get.results.heterosked(res,"sp.3.v2")),
  data.frame(mean="sp",  var="v3",get.results.heterosked(res,"sp.3.v3")),
  data.frame(mean="sp",  var="v4",get.results.heterosked(res,"sp.3.v4")),
  data.frame(mean="sp",  var="v5",get.results.heterosked(res,"sp.3.v5")),
  data.frame(mean="bump",var="v2",get.results.heterosked(res,"bump.3.v2")),
  data.frame(mean="bump",var="v3",get.results.heterosked(res,"bump.3.v3")),
  data.frame(mean="bump",var="v4",get.results.heterosked(res,"bump.3.v4")),
  data.frame(mean="bump",var="v5",get.results.heterosked(res,"bump.3.v5")),
  data.frame(mean="blk", var="v2",get.results.heterosked(res,"blk.3.v2")),
  data.frame(mean="blk", var="v3",get.results.heterosked(res,"blk.3.v3")),
  data.frame(mean="blk", var="v4",get.results.heterosked(res,"blk.3.v4")),
  data.frame(mean="blk", var="v5",get.results.heterosked(res,"blk.3.v5")),
  data.frame(mean="ang", var="v2",get.results.heterosked(res,"ang.3.v2")),
  data.frame(mean="ang", var="v3",get.results.heterosked(res,"ang.3.v3")),
  data.frame(mean="ang", var="v4",get.results.heterosked(res,"ang.3.v4")),
  data.frame(mean="ang", var="v5",get.results.heterosked(res,"ang.3.v5")),
  data.frame(mean="dop", var="v2",get.results.heterosked(res,"dop.3.v2")),
  data.frame(mean="dop", var="v3",get.results.heterosked(res,"dop.3.v3")),
  data.frame(mean="dop", var="v4",get.results.heterosked(res,"dop.3.v4")),
  data.frame(mean="dop", var="v5",get.results.heterosked(res,"dop.3.v5")),
  data.frame(mean="blip",var="v2",get.results.heterosked(res,"blip.3.v2")),
  data.frame(mean="blip",var="v3",get.results.heterosked(res,"blip.3.v3")),
  data.frame(mean="blip",var="v4",get.results.heterosked(res,"blip.3.v4")),
  data.frame(mean="blip",var="v5",get.results.heterosked(res,"blip.3.v5")),
  data.frame(mean="cor", var="v2",get.results.heterosked(res,"cor.3.v2")),
  data.frame(mean="cor", var="v3",get.results.heterosked(res,"cor.3.v3")),
  data.frame(mean="cor", var="v4",get.results.heterosked(res,"cor.3.v4")),
  data.frame(mean="cor", var="v5",get.results.heterosked(res,"cor.3.v5"))))
```